### 2022-03-03 doc/notebook/notebook.Rmd

---
title: "Notebook"
author: "Roland Knapp"
date: "`r Sys.Date()`"
editor_options:
  markdown:
    mode: gfm
---

## Predictors of frog survival following translocation

### Dataset structure

-   site_id: site to which frogs were translocated
-   elevation
-   release_date
-   release_day
-   release_year
-   siteid_donor
-   order: first translocation to site = 0, subsequent translocation = 1
-   type: translocation conducted on foot versus by helicopter
-   shore: suitability of shore/bank habitat for frog overwintering
-   pit_tag_ref
-   survival: individual-level estimated survival 1 year following translocation, based on cmr surveys conducted for at least two years post-translocation
-   sex
-   length: at release
-   weight: at release
-   condition: index based on length and weight
-   swab_id
-   bd_load: at release

Maximum depth and surface area would seem useful to include, but 74976 is a stream/meadow habitat and as such depth and area at that site are likely not comparable to depth and area at lake habitats.

### Model structure

survival \~ elevation + siteid_donor + release_day + release_year + order + shore + sex + length + condition + log(bdload) \| site_id

Group-level effects: In addition to the inclusion of site_id as a group-level effect, may need to include release_year, perhaps nested within site_id? Given that some sites only have a single release_year, will this cause problems?