### 2022-03-03 doc/manuscript/pnas_demo/pnas_demo.Rmd
Add PNAS manuscript template

---
title: Reintroduction of disease-resistant individuals allows recovery of an endangered frog in the presence of a lethal fungal disease

# Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
author:
  - name: Roland A. Knapp
    affiliation: a,b,1
  - name: Thomas C. Smith
    affiliation: a,b
  - name: Mark Q. Wilber
    affiliation: c
  - name: Andrew P. Rothstein
    affiliation: d,e,2
  - name: Allison Q. Byrne
    affiliation: d,e,2
  - name: Maxwell B. Joseph
    affiliation: f

address:
  - code: a
    address: Sierra Nevada Aquatic Research Laboratory, University of California, Mammoth Lakes, CA 93546
  - code: b
    address: Earth Research Institute, University of California-Santa Barbara, Santa Barbara, CA 93106
  - code: c
    address: Department of Forestry, Wildlife, and Fisheries, University of Tennessee, Knoxville, TN, xxxxx
  - code: d
    address: Department of Environmental Science, Policy, and Management, University of California-Berkeley, Berkeley, CA, 94720
  - code: e
    address: Museum of Vertebrate Zoology, University of California-Berkeley, Berkeley, CA, 94720
  - code: f
    address: Earth Lab, University of Colorado, Boulder, CO, 80303

corresponding_author:
  code: 1
  text: "To whom correspondence should be addressed. E-mail: roland.knapp@ucsb.edu"

# For footer text
lead_author_surname: Knapp

## Remove this if not required    
equal_authors:
  code: 2
  text: "A.P.R. and A.Q.B. contributed equally to this work."

author_contributions: |
  R.A.K, T.C.S, M.Q.W, A.P.R., A.Q.B, and M.B.J. designed research; R.A.K, T.C.S, M.Q.W, A.P.R., and A.Q.B performed research; R.A.K, M.Q.W, A.P.R., A.Q.B, and M.B.J. analyzed data; M.B.J. contributed software; and R.A.K, T.C.S, M.Q.W, A.P.R., A.Q.B, and M.B.J. wrote the paper.

## Remove this if not required
conflict_of_interest: |
  The authors declare no conflicts of interest.

abstract: |
  Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.

significance: |
  Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality. The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership. The Significance Statement appears in the paper itself and is required for all research papers.

acknowledgements: |
  Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.

keywords:
  - population recovery
  - amphibian declines
  - emerging infectious disease
  - reintroduction
  - Batrachochytrium dendrobatis

## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
pnas_type: pnasresearcharticle

bibliography: pnas.bib
csl: pnas.csl

## change to true to add optional line numbering
lineno: false

output: rticles::pnas_article
---

This PNAS journal template is provided to help you write your work in the
correct journal format. Instructions for use are provided below.

Note: please start your introduction without including the word
"Introduction" as a section heading (except for math articles in the
Physical Sciences section); this heading is implied in the first
paragraphs.

Results {.unnumbered}
========================================

Evolution of Reduced Disease Susceptibility {#results-subheading1 .unnumbered}
-------------------

Reintroduction of Disease-resistant Frogs {#results-subheading2 .unnumbered}
-------------------

Viability of Reestablished Populations {#results-subheading3 .unnumbered}
-------------------

Conceptual Model of Population Establishment {#results-subheading4 .unnumbered}
-------------------

Discussion {.unnumbered}
========================================

Conclusions {#discussion-subheading1 .unnumbered}
-------------------

Materials and Methods {.unnumbered}
========================================

Statistical Analyses {#methods-subheading1 .unnumbered}
-------------------

Data Availability {#methods-subheading2 .unnumbered}
-------------------

Guide to using this template {.unnumbered}
========================================

Please note that whilst this template provides a preview of the typeset
manuscript for submission, to help in this preparation, it will not
necessarily be the final publication layout. For more detailed
information please see the [PNAS Information for
Authors](http://www.pnas.org/site/authors/format.xhtml).

Author Affiliations {#author-affiliations .unnumbered}
-------------------

Include department, institution, and complete address, with the
ZIP/postal code, for each author. Use lower case letters to match
authors with institutions, as shown in the example. Authors with an
ORCID ID may supply this information at submission.

Submitting Manuscripts {#submitting-manuscripts .unnumbered}
----------------------

All authors must submit their articles at
[PNAScentral](http://www.pnascentral.org/cgi-bin/main.plex). If you are
using Overleaf to write your article, you can use the "Submit to PNAS"
option in the top bar of the editor window.

Format {#format .unnumbered}
------

Many authors find it useful to organize their manuscripts with the
following order of sections; Title, Author Affiliation, Keywords,
Abstract, Significance Statement, Results, Discussion, Materials and
methods, Acknowledgments, and References. Other orders and headings are
permitted.

Manuscript Length {#manuscript-length .unnumbered}
-----------------

PNAS generally uses a two-column format averaging 67 characters,
including spaces, per line. The maximum length of a Direct Submission
research article is six pages and a PNAS PLUS research article is ten
pages including all text, spaces, and the number of characters displaced
by figures, tables, and equations. When submitting tables, figures,
and/or equations in addition to text, keep the text for your manuscript
under 39,000 characters (including spaces) for Direct Submissions and
72,000 characters (including spaces) for PNAS PLUS.

References {#references .unnumbered}
----------

References should be cited in numerical order as they appear in text;
this will be done automatically via bibtex, e.g. @belkin2002using and
@berard1994embedding [@coifman2005geometric]. All references, including
for the SI, should be included in the main manuscript file. References
appearing in both sections should not be duplicated. SI references
included in tables should be included with the main reference section.

Data Archival {#data-archival .unnumbered}
-------------

PNAS must be able to archive the data essential to a published article.
Where such archiving is not possible, deposition of data in public
databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata,
and others outlined in the Information for Authors, is acceptable.

Language-Editing Services {#language-editing-services .unnumbered}
-------------------------

Prior to submission, authors who believe their manuscripts would benefit
from professional editing are encouraged to use a language-editing
service (see list at www.pnas.org/site/authors/language-editing.xhtml).
PNAS does not take responsibility for or endorse these services, and
their use has no bearing on acceptance of a manuscript for publication.

![Placeholder image of a frog with a long example caption to show
justification setting.<span data-label="fig:frog"></span>](frog.png)

Digital Figures {#sec:figures .unnumbered}
---------------

Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for
figures that will appear in the main text, and images must be final
size. Authors may submit U3D or PRC files for 3D images; these must be
accompanied by 2D representations in TIFF, EPS, or high-resolution PDF
format. Color images must be in RGB (red, green, blue) mode. Include the
font files for any text.

Figures and Tables should be labelled and referenced in the standard way
using the `\label{}` and `\ref{}` commands.

Figure \[fig:frog\] shows an example of how to insert a column-wide
figure. To insert a figure wider than one column, please use the
`\begin{figure*}...\end{figure*}` environment. Figures wider than one
column should be sized to 11.4 cm or 17.8 cm wide.

Single column equations {#single-column-equations .unnumbered}
-----------------------

Authors may use 1- or 2-column equations in their article, according to
their preference.

To allow an equation to span both columns, options are to use the
`\begin{figure*}...\end{figure*}` environment mentioned above for
figures, or to use the `\begin{widetext}...\end{widetext}` environment
as shown in equation \[eqn:example\] below.

Please note that this option may run into problems with floats and
footnotes, as mentioned in the [cuted package
documentation](http://texdoc.net/pkg/cuted). In the case of problems
with footnotes, it may be possible to correct the situation using
commands `\footnotemark` and `\footnotetext`.

$$\begin{aligned}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \label{eqn:example} \\
       &=x^3+3x^2y+3xy^3+x^3. 
\end{aligned}$$

<!-- pandoc writes all tables using longtable, which fails in 2-column mode

  Species                    CBS     CV     G3
  ----------------------- ------ ------ ------
  1\. Acetaldehyde           0.0    0.0    0.0
  2\. Vinyl alcohol          9.1    9.6   13.5
  3\. Hydroxyethylidene     50.8   51.2   54.0

  : Comparison of the fitted potential energy surfaces and ab initio
  benchmark electronic energy calculations

-->

Supporting Information (SI) {#supporting-information-si .unnumbered}
---------------------------

The main text of the paper must stand on its own without the SI. Refer
to SI in the manuscript at an appropriate point in the text. Number
supporting figures and tables starting with S1, S2, etc. Authors are
limited to no more than 10 SI files, not including movie files. Authors
who place detailed materials and methods in SI must provide sufficient
detail in the main text methods to enable a reader to follow the logic
of the procedures and results and also must reference the online
methods. If a paper is fundamentally a study of a new method or
technique, then the methods must be described completely in the main
text. Because PNAS edits SI and composes it into a single PDF, authors
must provide the following file formats only.

### SI Text {#si-text .unnumbered}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a
PDF with the same file name for visual reference).

### SI Figures {#si-figures .unnumbered}

Provide a brief legend for each supporting figure after the supporting
text. Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or
GIF format; figures may not be embedded in manuscript text. When saving
TIFF files, use only LZW compression; do not use JPEG compression. Do
not save figure numbers, legends, or author names as part of the image.
Composite figures must be pre-assembled.

### 3D Figures {#d-figures .unnumbered}

Supply a composable U3D or PRC file so that it may be edited and
composed. Authors may submit a PDF file but please note it will be
published in raw format and will not be edited or composed.

### SI Tables {#si-tables .unnumbered}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a
PDF with the same file name for visual reference); include only one
table per file. Do not use tabs or spaces to separate columns in Word
tables.

### SI Datasets {#si-datasets .unnumbered}

Supply Excel (.xls), RTF, or PDF files. This file type will be published
in raw format and will not be edited or composed.

### SI Movies {#si-movies .unnumbered}

Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media
(wmv), animated GIF (gif), or MPEG files and submit a brief legend for
each movie in a Word or RTF file. All movies should be submitted at the
desired reproduction size and length. Movies should be no more than 10
MB in size.

### Still images {#still-images .unnumbered}

Authors must provide a still image from each video file. Supply TIFF,
EPS, high-resolution PDF, JPEG, or GIF files.

### Appendices {#appendices .unnumbered}

PNAS prefers that authors submit individual source files to ensure
readability. If this is not possible, supply a single PDF file that
contains all of the SI associated with the paper. This file type will be
published in raw format and will not be edited or composed.

<!-- Leave these lines as they are at the end of your .Rmd file to ensure placement of methods & acknowledgements sections before the references-->
\showmatmethods
\showacknow
\pnasbreak