### 2022-03-07 doc/notebook/notebook_translocation.Rmd
Edits related to switch to github_document format

---
title: "Notebook: Predictors of frog survival following translocation"
author: "Roland Knapp"
date: "`r Sys.Date()`"
output:
  github_document: 
    toc: true
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, results='hide')
```

```{r}
library(tidyverse)
```

## Dataset structure

-   site_id: site to which frogs were translocated
-   elevation
-   date
-   day
-   year
-   donor
-   order: first translocation to site = 0, subsequent translocation = 1
-   type: translocation conducted on foot versus by helicopter (not included)
-   shore: suitability of shore/bank habitat for frog overwintering
-   pit_tag_ref
-   survival: individual-level estimated survival 1 year following translocation, based on cmr surveys conducted for at least two years post-translocation
-   sex
-   length: at release
-   weight: at release
-   condition: index based on length and weight
-   swab_id
-   bd_load: at release

Maximum depth and surface area would seem useful to include, but 74976 is a stream/meadow habitat and as such depth and area at that site are likely not comparable to depth and area at lake habitats.

Frog condition might be problematic because size and residuals are correlated. Specifically, the largest frogs have the largest positive residuals from a weight:length regression line, even after log10 transformation. As such, length and condition are probably not independent.

```{r frog-length-weight-plot}
read_csv(here::here("data", "clean", "frog_translocation.csv")) %>% 
ggplot(aes(x = log10(length), y = log10(weight))) +
  geom_point() +
  geom_smooth(method = "lm")
plot
```

Another consideration when designing the analysis is that there are 12 records in which bd_load is null. Of these, 9 are from 70413 on 2013-07-15, likely due to a rush to get frogs packaged prior to helicopter arrival. 39 frogs were translocated on that date, so should probably drop the 9 records where bd_load = NA prior to running the analysis. If bd_load is not an important predictor, can run analysis using the full dataset.

## Model structure

survival \~ elevation + shore + day + year + order + donor + sex + length + log(bdload) \| site_id

Group-level effects: In addition to the inclusion of site_id as a group-level effect, may need to include release_year, perhaps nested within site_id? Given that some sites only have a single release_year, will this cause problems?

Final steps before building model (in analysis Rmd):

-   Change to factor: order, day?, year?
-   Standardize variables?
-   Survival: Create 0/1 variable from probabilities. Change to factor?

## Results

To be added soon.