### 2022-10-29 code/translocation_survival_analysis_suppl.Rmd
Add dataset creation for analysis of bd load changes

---
title: "Predictors of frog survival following translocations - supplemental analyses"
author: "Roland Knapp"
output: html_notebook
---

The supplemental analyses in this notebook provide additional details to analyses in `translocation_survival_analysis.Rmd`.

## Load packages

```{r load-packages}
library(tidyverse)
library(brms)
library(rstanarm)
library(bayesplot)
library(tidybayes)
library(loo)
library(broom.mixed)
library(janitor)

source("classification_summary.R")
```

## Evaluate effect of frog size independent of donor population: Model m1d from `translocation_survival_analysis.Rmd` but using data subset where donor = 72996

### Read in frog translocation dataset  
* Read in data, create variables for analysis, standardize continuous variables. 
* arm package used to rescale continuous variables (substract mean, divide by 2 SD), manual centering of binary variables.
```{r read-translocation-data}
d1 <- read_csv(here::here("data", "clean", "frog_translocation_final.csv")) %>% 
  drop_na(bd_load) %>%  # drop 12 records where bd_load is NA (9 are from 70413)
  mutate(
    first = if_else(order == 1, 1, 0),
    surv = if_else(survival < 0.5, 0, 1),
    male = if_else(sex == "m", 1, 0),
    elev_lo = if_else(elevation < 3020, 1, 0), # using interval defined by cut_interval, n = 2
    elev_z = arm::rescale(elevation),
    snowt_z = arm::rescale(snow_t),
    snowt1_z = arm::rescale(snow_t1),
    day_z = arm::rescale(day),
    length_z = arm::rescale(length),
    bdload_l = log10(bd_load + 1),
    bdload_z = arm::rescale(bdload_l),
    shore_c = shore - mean(shore),
    male_c = male - mean(male),
    elevlo_c = elev_lo - mean(elev_lo),
    first_c = first - mean(first),
    across(c(shore_c, male_c, elevlo_c, first_c), round, 3), 
    across(c(site_id, shore_c, first_c, donor, male_c, elevlo_c, year), as.factor),
    surv = as.integer(surv))

# Add a descriptive translocation_id
d1 <- d1 %>% 
  distinct(site_id, year) %>% 
  arrange(site_id, year) %>% 
  group_by(site_id) %>% 
  mutate(transno = seq_len(n())) %>% 
  ungroup(site_id) %>% 
  unite(translocation_id, c("site_id", "transno"), remove = FALSE) %>% 
  select(-transno) %>% 
  mutate(translocation_id = as.factor(translocation_id)) %>% 
  inner_join(d1, by = c("site_id", "year")) %>% 
  relocate(translocation_id, .after = site_id)
```

### Model specification
```{r}
d1_72996 <- d1 %>% 
  filter(donor == "72996" & !is.na(bd_load))

m1d_72996 <- stan_glmer(
  surv ~ length_z + snowt_z + snowt1_z + day_z + bdload_z + elev_z + first_c + male_c + (1 | translocation_id),
  data = d1_72996,
  family = binomial,
  chains = 4, 
  iter = 5000*2, 
  cores = 4,
  seed = 84735)
```

### Priors & chain diagnostics
```{r}
prior_summary(m1d_72996)

b1_72996 <- c("(Intercept)", "length_z", "snowt_z", "snowt1_z", "day_z", "bdload_z", "elev_z", "first_c0.353", "male_c0.548")

mcmc_trace(m1d_72996, size = 0.1, pars = b1_72996)
mcmc_dens_overlay(m1d_72996, pars = b1_72996)

summary(m1d_72996)
```

### Posterior predictive check
```{r}
pp_check(m1d_72996, plotfun = "stat") +
    xlab("Frog survival rate")
```

### Posterior analysis
```{r}
tidy(m1d_72996, conf.int = TRUE, conf.level = 0.90)

mcmc_areas(m1d_72996, area_method = "equal height", prob = 0.90, pars = b1_72996) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with standardized predictors & translocation_id as grouping variable",
          "Posterior distributions with medians & 90% uncertainty intervals")

mean(bayes_R2(m1d_72996))

# Frog survival by translocation_id
mcmc_areas_ridges(m1d_72996, regex_pars = "translocation_id") +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Frog survival for each translocation_id")
```
- Using dataset containing only frogs translocated from 72996 had relatively little effect on the predictors of frog survival. Specific to the objective of this analysis, frog size remained unimportant as a predictor of survival. However, effect of Bd load became marginally unimportant instead of unimportant (i.e., uncertainty interval includes 0 but barely).

## Assess effect on m1d of including all records, including where bd_load = NA

### Read in frog translocation dataset  
* Read in data, create variables for analysis, standardize continuous variables. 
* arm package used to rescale continuous variables (substract mean, divide by 2 SD), manual centering of binary variables.
```{r read-translocation-data-including-na}
d1_allna <- read_csv(here::here("data", "clean", "frog_translocation_final.csv")) %>% 
#  drop_na(bd_load) %>%  # drop 12 records where bd_load is NA (9 are from 70413)
  mutate(
    first = if_else(order == 1, 1, 0),
    surv = if_else(survival < 0.5, 0, 1),
    male = if_else(sex == "m", 1, 0),
    elev_lo = if_else(elevation < 3020, 1, 0), # using interval defined by cut_interval, n = 2
    elev_z = arm::rescale(elevation),
    snowt_z = arm::rescale(snow_t),
    snowt1_z = arm::rescale(snow_t1),
    day_z = arm::rescale(day),
    length_z = arm::rescale(length),
    bdload_l = log10(bd_load + 1),
    bdload_z = arm::rescale(bdload_l),
    shore_c = shore - mean(shore),
    male_c = male - mean(male),
    elevlo_c = elev_lo - mean(elev_lo),
    first_c = first - mean(first),
    across(c(shore_c, male_c, elevlo_c, first_c), round, 3), 
    across(c(site_id, shore_c, first_c, donor, male_c, elevlo_c, year), as.factor),
    surv = as.integer(surv))

# Add a descriptive translocation_id
d1_allna <- d1_allna %>% 
  distinct(site_id, year) %>% 
  arrange(site_id, year) %>% 
  group_by(site_id) %>% 
  mutate(transno = seq_len(n())) %>% 
  ungroup(site_id) %>% 
  unite(translocation_id, c("site_id", "transno"), remove = FALSE) %>% 
  select(-transno) %>% 
  mutate(translocation_id = as.factor(translocation_id)) %>% 
  inner_join(d1_allna, by = c("site_id", "year")) %>% 
  relocate(translocation_id, .after = site_id)
```

### Model specification
```{r}
m1d_allna <- stan_glmer(
  surv ~ length_z + snowt_z + snowt1_z + day_z + elev_z + first_c + male_c + donor + (1 | translocation_id),
  data = d1_allna,
  family = binomial,
  chains = 4, 
  iter = 5000*2, 
  cores = 4,
  seed = 84735)
```

### List model variable names
```{r}
get_variables(m1d_allna)
```

### Priors & chain diagnostics
```{r}
prior_summary(m1d_allna)

b1_allna <- c("(Intercept)", "length_z", "snowt_z", "snowt1_z", "day_z", "elev_z", "first_c0.349", "male_c0.549", "donor70567", "donor72996")

mcmc_trace(m1d_allna, size = 0.1, pars = b1_allna)
mcmc_dens_overlay(m1d_allna, pars = b1_allna)

summary(m1d_allna)
```

### Posterior predictive check
```{r}
pp_check(m1d_allna, plotfun = "stat") +
    xlab("Frog survival rate")
```

### Posterior analysis
```{r}
tidy(m1d_allna, conf.int = TRUE, conf.level = 0.90)

mcmc_areas(m1d_allna, area_method = "equal height", prob = 0.90, pars = b1_allna) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with standardized predictors & translocation_id as grouping variable",
          "Posterior distributions with medians & 90% uncertainty intervals")

mean(bayes_R2(m1d_allna))

# Frog survival by translocation_id
mcmc_areas_ridges(m1d_allna, regex_pars = "translocation_id") +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Frog survival for each translocation_id")
```

- Inclusion of all records (i.e., 12 records where bd_load = NA that were excluded in primary analysis) did not change the results.

## Analysis of changes in Bd load following translocation

-  Rescale bd_load, center period variable. 
-  Compare loads at time of translocation to those from CMR years 0 and 1. Exclude years > 1.
-  Possible models: 
  -  bd_load ~ period + (1 | pit_tag_ref) + (1 | translocation_id)
  -  bd_load ~ period + (1 | pit_tag_ref) + (1 | site_id)

### Load additional libraries
```{r}
library(RPostgreSQL)
library(lubridate)
```

### Read in raw translocation and capture datasets, bind them
```{r read-data}
translocation <- read_csv(here::here("data", "raw", "frog_translocation.csv")) %>% 
  relocate(visit_date = release_date, .before = type) %>% 
  rename(survey_type = type) %>% 
  mutate(pit_tag_ref = replace(pit_tag_ref, pit_tag_ref == "900067000117396" & site_id == 70641, "111111111111111")) %>% # remove incorrect pit_tag_ref identified in translocation_survival_createdata.Rmd
  filter(pit_tag_ref != "111111111111111")

capture <- read_csv(here::here("data", "raw", "frog_captures.csv")) %>% 
  select(-capture_life_stage, -capture_animal_state, -sex)

# select only translocated frogs (exclude naturally-recruited frogs and translocated frogs that were never recaptured after translocation)
capture <- translocation %>% 
  select(pit_tag_ref) %>% 
  inner_join(capture, by = c("pit_tag_ref")) %>% 
  relocate(pit_tag_ref, .after = last_col())

frogs_all <- bind_rows(translocation, capture)
```

### Connect to PostgreSQL database `amphibians`
```{r}
con = dbConnect(dbDriver("PostgreSQL"),
                user = rstudioapi::askForPassword("user name"),
                password = rstudioapi::askForPassword("password"),
                host = "frogdb.eri.ucsb.edu",
                port = 5432,
                dbname = "amphibians")

```

### Retrieve Bd data for translocated frogs from database
```{sql retrieve-translocation-bd-data, connection=con, output.var="translocate_bd"}
select
  release_date as visit_date,
  pit_tag_ref,
  swab_id,
  bd_load
from relocate
inner join relocate_frog on relocate.id = relocate_id
inner join bd_load on swab_id = sample_id  
where release_siteid1 in (70134, 70279, 70370, 70413, 70414, 70449, 70505, 70550, 70556, 70619, 70628, 70641, 74976) and
  (replicate = 1 or replicate is null)
```

### Retrieve Bd data for captured frogs from database
```{sql retrieve-capture-bd-data, connection=con, output.var="capture_bd"}
select
	visit_date,
	pit_tag_ref, 
	swab_id,
	bd_load
from visit
inner join survey on visit.id = visit_id
inner join capture_survey on survey.id = survey_id
inner join bd_load on swab_id = sample_id  
where site_id in (70134, 70688, 70370, 70114, 70175, 70279, 70413, 71570, 71968, 72008, 72093, 72264, 72390, 72442, 72694, 70414, 70449, 70034, 70505, 72092, 70550, 70556, 70619, 70628, 70641, 74976, 72973) and 
	survey_type = 'cmr' and
	capture_animal_state != 'dead' and
	(replicate = 1 or replicate is null)
```

### Disconnect from database
```{r}
dbDisconnect(con) 
rm("con")
```

### Combine Bd data from translocate and capture datasets
```{r}
bd_add <- bind_rows(translocate_bd, capture_bd)
```

### Add Bd data to frog data, restructure dataset
```{r}
frogsbd <- frogs_all %>% 
  inner_join(bd_add, by = c("visit_date", "pit_tag_ref")) %>% 
  mutate(period = if_else(survey_type == "cmr", "after", "before")) %>% # use 0/1 instead?
  relocate(period, .after = survey_type) %>% 
  arrange(site_id, visit_date)
```

### Add years since translocation
```{r}
frogsbd <- translocation %>% 
  mutate(year_translocate = year(visit_date)) %>% 
  select(year_translocate, pit_tag_ref) %>% 
  inner_join(frogsbd, by = c("pit_tag_ref")) %>% 
  relocate(year_translocate, .after = visit_date) %>% 
  relocate(pit_tag_ref, .after = period) %>% 
  mutate(year_capture = year(visit_date), .after = year_translocate) %>% 
  mutate(interval = (year_capture - year_translocate), .after = year_capture) %>% 
  arrange(site_id, visit_date, pit_tag_ref)
```

### Update site ids to standardize usage
```{r}
frogsbd <- frogsbd %>% 
  mutate(site_id1 = site_id, .after = site_id,
         site_id1 = replace(site_id1, site_id == 70034, 70449),
         site_id1 = replace(site_id1, site_id == 70175, 70413),
         site_id1 = replace(site_id1, site_id == 70279, 70413),
         site_id1 = replace(site_id1, site_id == 72093, 70413),
         site_id1 = replace(site_id1, site_id == 72442, 70413),
         site_id1 = replace(site_id1, site_id == 70688, 70134),
         site_id1 = replace(site_id1, site_id == 72092, 70505),
         site_id1 = replace(site_id1, site_id == 72973, 74976))
```

### Add translocation_id
```{r}
frogsbd <- frogsbd %>% 
  distinct(site_id1, year_translocate) %>% 
  arrange(site_id1, year_translocate) %>% 
  group_by(site_id1) %>% 
  mutate(transno = seq_len(n())) %>% 
  ungroup(site_id1) %>% 
  unite(translocation_id, c("site_id1", "transno"), remove = FALSE) %>% 
  select(-transno) %>% 
  inner_join(frogsbd, by = c("site_id1", "year_translocate")) %>% 
  select(site_id, site_id1, translocation_id, visit_date, survey_type, period, year_translocate, year_capture, interval, pit_tag_ref, swab_id, bd_load) %>% 
  arrange(site_id1, visit_date)
```

### Save frog-Bd dataset
```{r}
frogsbd %>% write_csv(here::here("data", "clean", "bd_beforeafter_translocation.csv"))
```