### 2022-11-23 code/maps.qmd
Add inset map to map of study area

---
title: "maps"
format: html
editor: visual
---

## Load packages

```{r}
library(sf)
library(tmap)
library(stars)
library(grid)
library(tidyverse)
library(RPostgreSQL)
```

## Map of sample sites for exome capture analysis

Run the following 5 chunks only if you have access to the project database. For those without the necessary credentials, use the saved output file in data/maps/ that is available via the "import-map-data" chunk.

### Create dataset of sample collection points

```{r}
exome_sites <- read_csv(here::here("data", "maps", "exome_sample_sites.csv"))
```

#### Connect to database

```{r}
source("db_connect.R")
```

#### Retrieve data

```{sql retrieve-site-xy, connection=con, output.var="exome_sites_db"}
select
  id as site_id,
  utme,
  utmn
from site
where id in (74281, 50731, 50307, 52127, 10220, 10490, 50837, 20169, 72996, 72808, 70567, 54188, 10055, 10593, 50783)
```

#### Disconnect from database

```{r}
source("db_disconnect.R")
```

#### Create points layer

```{r}
exome_sites_xy <- 
  exome_sites %>% 
  inner_join(exome_sites_db, by = "site_id")

# Add x-y random noise to obscure exact site locations
exome_sites_xy <- exome_sites_xy %>% 
  add_column(x = sample(-2000:2000, size = 15),
             y = sample(-2000:2000, size = 15))

exome_sites_xy <- exome_sites_xy %>% 
  mutate(utme_noise = utme + x,
         utmn_noise = utmn + y) %>% 
  select(-utme, -utmn, -x, -y)

# Save file 
write_csv(exome_sites_xy, here::here("data", "maps", "exome_sites_xy.csv"))
```

### Import map layers

```{r import-map-data}
# Read file, if lacking database access
exome_sites_xy <- read_csv(here::here("data", "maps", "exome_sites_xy.csv"))
exome_points = st_as_sf(exome_sites_xy, coords = c("utme_noise", "utmn_noise"), crs = st_crs(26911))

nps_boundaries <- st_read(here::here(
  "data", "maps", "nps-boundaries", "yose_kica_sequ_boundaries.shp"))
st_crs(nps_boundaries) # determine coordinate reference system (CRS)
nps_boundaries <- st_transform(nps_boundaries, 26911) # transform to NAD83 UTM11

# California Color Hillshade (90m) from koordinates.com
hillshade_90m <- read_stars(here::here(
  "data", "maps", "hillshade-90m", "california-color-hillshade-90m.tif"))
st_crs(hillshade_90m)

hillshade_300m <- read_stars(here::here(
  "data", "maps", "hillshade-300m", "california-color-hillshade-300m.tif"))
```

### Check & fix shapefile polygons

```{r}
st_is_valid(nps_boundaries) # all 3 are invalid

nps_boundaries <- st_make_valid(nps_boundaries$geometry)

st_is_valid(nps_boundaries) # fix polygons
```

### Crop raster with bounding box

```{r}
bbox <- st_bbox(c(xmin = 238000, ymax = 4235000, xmax = 403000, ymin = 4011000), 
                crs = st_crs(nps_boundaries)) %>%  
  st_as_sfc()
hillshade_crop <- st_crop(hillshade_90m, bbox)
hillshade_crop <- st_transform(hillshade_crop, 26911)
```

### Create study area map

```{r}
exome_points_map <- tm_shape(hillshade_crop) + 
  tm_rgb(r = 1, g = 2, b = 3) +
  tm_shape(nps_boundaries) + 
  tm_borders(col = "black") +
  tm_shape(exome_points) + 
  tm_symbols(size = 0.7, # increase point size, use darker blue
             col = "bd_status", 
             palette = c("#FFEC8B", "#00B2EE"), # tmap only recognizes basic color names (e.g., "blue" but not "lightblue") - use HEX instead
             title.col = "Bd status") +
  tm_layout(legend.position = c("right", "top")) +
  tm_compass(type = "4star",
             position = c("left", "bottom"),
             size = 2) +
  tm_scale_bar(breaks = c(0, 25, 50), 
               text.size = 0.7, 
               position = c("left", "bottom"))

exome_points_map
```

### Create inset map of California

```{r}
inset_map <- tm_shape(hillshade_300m) +
  tm_rgb(r = 1, g = 2, b = 3) +
  tm_shape(bbox) + 
  tm_borders(col = "black")

inset_map
```

### Study area map with inset map

```{r}
exome_points_map
print(inset_map, vp = viewport(0.35, 0.37, width = 0.2, height = 0.2))
```

### Save map

```{r}
tmap_save(exome_points_map, 
          here::here("out", "map_exome_points.png"),
          insets_tm = inset_map,
          insets_vp = viewport(0.14, 0.24, width = 0.2, height = 0.2))
```

#### To Do

-   Exported CA tif has no whitespace around state, has black border. How to remove border and add whitespace?