### 2023-05-05 doc/manuscript/PNAS-template-empty.tex
Convert qmd to latex PNAS format

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

\title{}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,b]{Roland A. Knapp}
\author[c,1]{Mark Q. Wilber} 
\author[d,e,1]{Allison Q. Byrne}
\author[f,g]{Maxwell B. Joseph}
\author[a,b]{Thomas C. Smith}
\author[d,e]{Andrew P. Rothstein}
\author[h]{Robert L. Grasso}
\author[d,e]{Erica Bree Rosenblum}

\affil[a]{Sierra Nevada Aquatic Research Laboratory, University of California, Mammoth Lakes, CA, 93546}
\affil[b]{Earth Research Institute, University of California, Santa Barbara, CA, 93106-3060}
\affil[c]{School of Natural Resources, University of Tennessee Institute of Agriculture, Knoxville, TN, 37996}
\affil[d]{Department of Environmental Science, Policy, and Management, University of California - Berkeley, Berkeley, CA, 94720-3114}
\affil[e]{Museum of Vertebrate Zoology, University of California - Berkeley, Berkeley, CA, 94720-3160}
\affil[f]{Earth Lab, University of Colorado, Boulder, CO, 80303}
\affil[g]{Planet, San Francisco, CA, 94107}
\affil[h]{Resources Management and Science, Yosemite National Park, El Portal, CA, 95318}

% Please give the surname of the lead author for the running footer
\leadauthor{Knapp}

% Please add a significance statement to explain the relevance of your work
\significancestatement{}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{}
\authordeclaration{}
\equalauthors{\textsuperscript{1}M.W. contributed equally to this work with A.B.}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: roland.knapp@ucsb.edu}

% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
\keywords{}

\begin{abstract}
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% Introduction

% Results

% Discussion

\matmethods{}

\showmatmethods{} % Display the Materials and Methods section

\acknow{}

\showacknow{} % Display the acknowledgments section

% Bibliography
\bibliography{translocation}

% Figures

\end{document}