### 2023-05-15 doc/manuscript/PNAS-template-supp-info-empty.tex

\documentclass[9pt,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnassupportinginfo}
\usepackage{longtable}

\title{}
\author{Roland A. Knapp, Mark Q. Wilber, Allison Q. Byrne, Maxwell B. Joseph, Thomas C. Smith, Andrew P. Rothstein, Robert L. Grasso, Erica Bree Rosenblum}
\correspondingauthor{Corresponding Author name. Roland A. Knapp \\ E-mail: roland.knapp@ucsb.edu}

\begin{document}

\maketitle

%% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
\SItext

% Figures

%%% Add this line AFTER all your figures and tables
\FloatBarrier

\dataset{gemma\_outliers\_all.csv}{Set of liberal SNP and INDEL outlier variants (Bonferroni corrected p-value $<$ 0.05), identified via GEMMA.}

\dataset{gemma\_outliers\_strict\_freq.csv}{Set of strict SNP and INDEL outlier variants (Bonferroni corrected p-value $<$ 0.01), identified via GEMMA.
Additional information includes variant location within the gene (predicted\_gene\_loc) and whether the variant is synonymous or nonsynonymous (predicted\_effect\_AA).}

\dataset{spline\_window\_shared\_outliers.csv}{Description of overlapping \emph{F\textsubscript{ST}} and
\(\pi_{diff}\) outlier windows, as identified in the splined window
analysis.}

\dataset{spline\_window\_gene\_details.csv}{Annotation information for all genes within each of the
overlapping outlier windows in dataset S3}

\bibliography{translocation}

\end{document}