### 2023-05-19 doc/manuscript/translocation_bioarxiv.tex
Address formatting comments on compiled manuscript

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasmathematics} % Choose template 
% {pnasmathematics} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

\title{Evolutionary rescue and reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,b]{Roland A. Knapp}
\author[c,1]{Mark Q. Wilber} 
\author[d,e,1]{Allison Q. Byrne}
\author[f,g]{Maxwell B. Joseph}
\author[a,b]{Thomas C. Smith}
\author[d,e,h]{Andrew P. Rothstein}
\author[i]{Robert L. Grasso}
\author[d,e]{Erica Bree Rosenblum}

\affil[a]{Sierra Nevada Aquatic Research Laboratory, University of California, Mammoth Lakes, CA, 93546}
\affil[b]{Earth Research Institute, University of California, Santa Barbara, CA, 93106-3060}
\affil[c]{School of Natural Resources, University of Tennessee Institute of Agriculture, Knoxville, TN, 37996}
\affil[d]{Department of Environmental Science, Policy, and Management, University of California - Berkeley, Berkeley, CA, 94720-3114}
\affil[e]{Museum of Vertebrate Zoology, University of California - Berkeley, Berkeley, CA, 94720-3160}
\affil[f]{Earth Lab, University of Colorado, Boulder, CO, 80303}
\affil[g]{Current affiliation: Planet, San Francisco, CA, 94107}
\affil[h]{Current affiliation: Ginkgo Bioworks, Boston, MA, 02210}
\affil[i]{Resources Management and Science, Yosemite National Park, El Portal, CA, 95318}

% Please give the surname of the lead author for the running footer
\leadauthor{Knapp}

% Please add a significance statement to explain the relevance of your work
% \significancestatement{Understanding how species persist despite accelerating global change is
% critical for the conservation of biodiversity. One possibility is rapid
% evolution that restores positive growth to declining populations
% (``evolutionary rescue''). Individuals from these rescued populations
% could also be used to reestablish extinct populations. We evaluated both
% possibilities in a once-common frog species driven to the brink of
% extinction by a disease that has devastated global amphibian
% biodiversity and for which few options exist to reverse disease impacts.
% We found that frogs have evolved increased resistance to infection,
% facilitating population recovery. In addition, introduction of frogs
% from these populations allowed reestablishment despite the ongoing
% presence of disease. Such reintroductions could play an important role
% in biodiversity conservation in our rapidly changing world.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{}
\authordeclaration{}
\equalauthors{\textsuperscript{1}M.W. contributed equally to this work with A.B.}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: roland.knapp@ucsb.edu}

% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
\keywords{}

\begin{abstract}
Vast alteration of the biosphere by humans is causing a sixth mass
extinction. Adaptation to modified environments often is the only means
for species to persist, but the extent to which such ``evolutionary
rescue'' can prevent or reverse biodiversity loss is largely unknown.
Using results from genomic analyses, a 15-year reintroduction effort,
and population modeling, we provide a compelling example of evolution
reversing the declines of an imperiled amphibian and allowing the
reestablishment of extirpated populations. The once-common mountain
yellow-legged (MYL) frog is threatened with extinction by the
human-facilitated emergence of a lethal fungal pathogen
(\emph{Batrachochytrium dendrobatidis}; ``Bd''). Although most MYL frog
populations are extirpated following disease outbreaks, some persist and
eventually begin to recover. Using an exome capture approach, we show
that MYL frogs have undergone substantial evolutionary change following
disease outbreaks. Moreover, in naturally recovering populations
observed changes in immune function loci are associated with increased
resistance/tolerance to Bd infection. Large-scale reintroduction of
frogs from rescued populations resulted in the establishment of
reproducing populations despite ongoing disease. In addition, results
from viability modeling suggest that established populations have a low
probability of extinction over 50 years. Collectively, these results
provide a rare example of how evolutionary rescue and the reintroduction
of resistant/tolerant individuals can allow the landscape-scale recovery
of disease-impacted species. This example has potentially broad
implications for the many taxa worldwide that are threatened with
extinction by anthropogenic stressors.
\end{abstract}

\dates{This manuscript was compiled on \today}
% \doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\dropcap{H}uman activities are intensifying threats to global biodiversity
\citep{ceballos2015}, with important implications for ecosystem
resilience and human welfare \citep{naeem2009}. To persist, affected
taxa will often need to adapt to their altered environment or face
decline and possible extinction. The process whereby rapid evolutionary
change increases the frequency of adaptive alleles and restores positive
population growth is known as ``evolutionary rescue''
\citep{carlson2014}. In addition to preventing extinction of declining
populations, individuals from rescued populations could also disperse,
or be reintroduced, to empty habitats and promote larger-scale recovery.
Although this overall scenario is potentially of critical importance for
the conservation of an ever-growing number of imperiled taxa,
descriptions of its occurrence and effectiveness in the wild are
lacking.

One consequence of human alteration of the biosphere is an increase in
emerging infectious diseases \citep{jones2008, fisher2012}. Such
diseases pose a severe threat to wildlife populations
\citep{daszak2000}, and have caused dramatic declines and extinctions in
a wide range of taxa, including echinoderms, mammals, birds, and
amphibians \citep{hewson2014, samuel2015, scheele2019, cunningham2021}.
Amid this global loss of biodiversity, some populations decimated by
disease have subsequently recovered to varying extents
\citep{scheele2017, voyles2018, knapp2016}. Although we know relatively
little about the mechanisms underlying recovery
\citep{brannelly2021, russell2020}, host evolution is one possibility
\citep{carlson2014, searle2020}. This process is known from a diversity
of wildlife populations
\citep{savage2016, epstein2016, gignoux-wolfsohn2021, holland2022}, and
often results in increased host resistance and/or tolerance \citep[see][
for definitions]{medzhitov2012}. However, linking these evolutionary
responses to population recoveries is difficult and few clear examples
exist. In addition to promoting recovery of disease-impacted
populations, the resistant/tolerant individuals in such populations
could potentially be used to reestablish populations that were
extirpated following disease emergence
\citep{mendelson2019, brannelly2021}. To date, the use of
resistant/tolerant individuals for species recovery is rare \citep[but
see][]{joseph2018}, and studies that evaluate its potential utility are
critically needed.

The mountain yellow-legged (MYL) frog, composed of the sister species
\emph{Rana muscosa} and \emph{Rana sierrae} \citep{vredenburg2007}, is
emblematic of the unprecedented global declines of amphibians caused by
the recently-emerged amphibian chytrid fungus
\citep[\emph{Batrachochytrium dendrobatidis,}][]{scheele2019}. Once the
most common amphibian in the high elevation portion of California's
Sierra Nevada mountains \citep[USA,][]{grinnell1924}, during the past
century this frog has disappeared from more than 90\% of its historical
range \citep{vredenburg2007}. Due to the severity of its decline and the
increasing probability of extinction, both species are now listed as
``endangered'' under the U.S. Endangered Species Act. In the Sierra
Nevada, this decline was initiated by the introduction of non-native
trout into the extensive historically-fishless region
\citep{bradford1989, knapp2000} starting in the late 1800s. The arrival
of \emph{B. dendrobatidis} (Bd) in the 1960s and its subsequent spread
\citep{vredenburg2019} caused additional large-scale population
extirpations \citep{vredenburg2010, rachowicz2006}. These Bd-caused
declines are fundamentally different from the fish-caused declines
because removal of fish populations is feasible and results in the rapid
recovery of frogs \citep{knapp2007, vredenburg2004}. In contrast, Bd
appears to persist in habitats even in the absence of amphibian hosts
\citep{walker2007}, and therefore represents a long-term alteration of
invaded ecosystems that amphibians will need to overcome to reestablish
populations.

Despite the catastrophic impact of Bd on MYL frogs, wherein most
Bd-naive populations are extirpated following Bd arrival
\citep{vredenburg2010}, some populations have persisted after epizootics
\citep{briggs2010} and are now recovering \citep{knapp2016}. Frogs in
these recovering populations show reduced susceptibility to Bd
infection, consistent with selection for more resistant/tolerant
genotypes \citep{knapp2016}. Furthermore, exploratory efforts to
reintroduce frogs collected from recovering populations to sites from
which frogs were previously extirpated indicate that population
establishment is possible despite ongoing Bd infection
\citep{joseph2018}.

In the current study, we combine results from genomic analyses, a
15-year landscape-scale frog reintroduction effort, and a detailed
population model to provide an unprecedented example of how evolution in
response to disease allows the successful reestablishment of extirpated
MYL frog populations. Specifically, we used exome capture methods to
show that MYL frogs have undergone substantial evolutionary change
associated with Bd exposure, changes that might provide a genetic basis
for the increased resistance/tolerance to Bd infection observed in
naturally recovering populations \citep{knapp2016}. We then use the
reintroduction of resistant/tolerant frogs to 24 sites and results from
long-term capture-mark-recapture (CMR) surveys to describe the dynamics
and drivers of population establishment. These data are also used to
parameterize a population model that extends our inferences of
population establishment well beyond the temporal extent of the CMR
data. Collectively, these results provide one of the first examples of
how evolutionary rescue and the reintroduction of resistant/tolerant
animals can allow the large-scale recovery of disease-impacted species.

\section*{Results}

\subsection*{Frog evolution in response to Bd}

To determine whether MYL frog populations show an evolutionary response
to Bd, we compared frog genomes between populations with contrasting
histories of Bd exposure. Specifically, we compared frog genomes sampled
in four populations that have not yet experienced a Bd-caused epizootic
(``naive'') \citep{zhou2015} versus in five populations that experienced
a Bd epizootic during the past several decades and have since recovered
to varying degrees (``recovering''; Figure~\ref{fig-allelefrequencies})
\citep{knapp2016, vredenburg2010}. These recovering populations exist in
an enzootic state \citep{briggs2010}, characterized by high Bd
prevalence and, in adults, moderate Bd infection intensity (``load'')
that is typically well below the level expected to cause mortality
\citep{vredenburg2010}. (See SI Appendix for additional details
regarding the study design.)

We conducted a principal component analysis (PCA) of the genomic data to
describe the relationships between sampled populations, and then used
two complementary approaches to identify regions of the genome under
selection. First, we used a multivariate linear mixed model to evaluate
associations between population type (i.e., naive versus recovering) and
individual variants, including single nucleotide polymorphisms (SNPs)
and insertions/deletions (INDELS). Second, we used a splined window
analysis to identify larger genomic regions showing differences between
population types in \emph{F\textsubscript{ST}} and nucleotide diversity
(\(\pi_{diff}\) = \(\pi_{naive}\) - \(\pi_{recovering}\)).

Individual frogs clustered into 3 separate groups in principal component
space (Figure S1A), and clusters reflected the
species split (i.e., \emph{R. muscosa} versus \emph{R. sierrae}) and the
strong signature of isolation by distance that is characteristic of MYL
frogs \citep{rothstein2020, poorten2017}. Importantly, each cluster
contained at least one population from both the naive and recovering
groups, allowing us to distinguish allelic associations of individuals
sampled in the 2 population types versus allelic associations resulting
from population structure and genetic drift.

Results from the individual variant and splined window analyses show
that recovering populations have signatures of strong selection on
multiple regions of the genome. The analysis of individual variants
identified 11 ``outlier'' SNPs (i.e., showing significantly different
allele frequencies between naive versus recovering populations) from 7
distinct genes across 4 contigs (Figure S1B, C;
SI Appendix: Datasets S1, S2). One of the 7 identified genes
(LOC108802036) did not have an associated annotation. For the outlier
SNPs, frequency differences between the naive and recovering populations
ranged from 41\% to 86\%. Most of these SNPs showed frequency
differences in only a subset of the sampled populations
(Figure~\ref{fig-allelefrequencies}A, B), but the SNP in the RIN3 gene
showed consistent differences in frequencies across all populations
(Figure~\ref{fig-allelefrequencies}C). This is suggestive of strong -
and parallel - selection at this locus associated with exposure of frogs
to Bd across multiple populations.

The splined window analysis identified 33 outlier regions for
\(\pi_{diff}\) and 58 outlier regions for \emph{F\textsubscript{ST}}
(Figure~\ref{fig-spline-manhattan}A, B). Of these, 9 regions were
outliers for both metrics (``shared regions'') and 2 of these shared
regions also contained one or more of the outlier SNPs described above.
A total of 33 annotated genes were found in the 9 shared regions (SI
Appendix: Datasets S3, S4). Given this large number of genes, here we
focus on those with the strongest signal of selection and/or
immune-related functions. The largest \(\pi_{diff}\), indicative of
directional selection, occurred in a 163kb region on Contig19, 12.9Mb
upstream of the RIN3 outlier SNP (Figure~\ref{fig-spline-manhattan}C).
This region contains approximately 500 SNPs and one annotated gene
called ``interferon-induced very large GTPase 1-like'' (GVINP1).
Additionally, a shared outlier region on Contig1 contained two
complement factor genes (C6 and C7). Interestingly, this region had a
large negative \(\pi_{diff}\) value, consistent with balancing
selection. Finally, one shared outlier region on Contig8 contained one
outlier SNP (TCF19) and was within 360kb of another outlier SNP (VARS)
(Figure~\ref{fig-spline-manhattan}D, Figure S2).
This region (854kb from the beginning of the outlier window to the VARS
SNP) contained a total of 8 annotated genes. In \emph{Xenopus}, five of
these genes occur in the extended major histocompatibility complex (MHC)
Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the
extended MHC Class III region (HSP70, LSM2, VARS) \citep{ohta2006}.
Therefore, this region under selection is part of the extended MHC Class
I and III complex and shows synteny with other amphibian genomes.

\subsection*{Frog population recovery}

To determine whether frogs from recovering populations could be used to
reestablish extirpated populations, we conducted 24 reintroductions in
Yosemite National Park (2006--2020). Each of the reintroductions
involved collection of adult frogs from 1 of 3 recovering, Bd-positive
``donor'' populations and translocating them \citep{seddon2014} to 1 of
12 nearby recipient sites (Figure S3). The donor
populations included 2 of the 5 recovering populations in the frog
evolution study described above (Figure~\ref{fig-allelefrequencies}A:
population 1 and 4). Following translocation, adult survival and
recruitment of new adults was estimated from CMR surveys and counts of
tadpoles and juveniles were obtained from visual encounter surveys
(VES).

Estimates of frog survival during the 1-year period following
translocation indicate that survival was highly variable between
recipient sites, but relatively constant within recipient sites (for the
subset of sites that received multiple translocations;
Figure~\ref{fig-translocation-survival}). As such, the average survival
of frogs in the first translocation to a site was an important predictor
of survival of frogs in subsequent translocations to that site
(Figure S4). This suggests an important
effect of site characteristics and/or (site * frog) interactions on frog
survival. In addition, 1-year survival was higher for frogs translocated
later in the study period than earlier
(Table S1), which we suggest resulted
primarily from our improved ability to choose recipient sites with
higher habitat quality for \emph{R. sierrae}. Finally, Bd loads were
fairly consistent before versus after translocation and loads were
nearly always well below the level indicative of severe chytridiomycosis
and associated frog mortality (Figure S5)
\citep{joseph2018, vredenburg2010}. Additional details on methods used
to estimate Bd loads are provided in the SI Appendix.

Of the 12 translocated populations, 9 (0.75) showed evidence of
successful reproduction in subsequent years, as indicated by the
presence of tadpoles and/or juveniles. For these 9 populations, one or
both life stages were detected in nearly all survey-years following
translocation (proportion of survey-years: median = 0.9, range =
0.29--1). These same populations were also those in which recruitment of
new adults (i.e., progeny of translocated individuals) was detected. As
with early life stages, recruits were detected in the majority of
post-translocation survey-years (proportion of survey-years: median =
0.79, range = 0.12--1). In summary, survey results indicate that
translocations resulted in the establishment of reproducing MYL frog
populations at most recipient sites despite the ongoing presence of Bd.

To determine whether Bd load had a negative effect on adult survival, as
would be expected if frogs were highly susceptible to Bd infection, we
developed linear and multilevel models. The best model of 1-year frog
survival identified several important predictors, but Bd load at the
time of translocation was not among them
(Figure S6). Important predictors included
winter severity in the year following translocation, site elevation, and
donor population (Figure~\ref{fig-cond-effects}). Males had somewhat
higher survival than females, but this effect was only marginally
important (Figure~\ref{fig-cond-effects},
Figure S6). Collectively, these results are
consistent with frogs in recovering populations having evolved
sufficient resistance/tolerance to minimize the adverse effects of Bd
infection. In addition, the moderate Bd loads that characterized frogs
following translocation (Figure S5) and
relatively high survival of frogs in numerous translocated cohorts
(Figure~\ref{fig-translocation-survival}) suggests that frogs
translocated from recovering populations can maintain the benefits of
resistance/tolerance in non-natal habitats. Finally, in 3 locations
where longer CMR time series allowed us to assess the survival of new
adults recruited to the population, naturally recruited adults had
equivalent or higher survival probabilities than the originally
translocated adults (Figure S7). Therefore,
frog resistance/tolerance in non-natal habitats is maintained across
generations.

\subsection*{Long-term population viability}

Results from frog translocations indicated that adult survival was often
relatively high and that most populations showed evidence of successful
reproduction and recruitment (described above). Although suggestive of
population establishment, a decade or more of surveys may be necessary
to confirm that populations are in fact self-sustaining
\citep{joseph2018}. To extend our inferences of population establishment
beyond those possible from the 2-16 years of site-specific CMR data, we
developed a population viability model. Specifically, to test whether
the observed yearly adult survival probabilities in translocated
populations were sufficient to promote long-term viability, we built a
stage-structured matrix model that captured known frog demography and
included demographic and environmental stochasticity. We parameterized
the model using CMR data from translocated populations and known life
history values in this system (Table S2).

Given observed yearly adult survival probabilities (from analysis of CMR
data) and a baseline yearly recruitment probability \(\omega\) of 0.3
(i.e., the probability that, following survival, juveniles successfully
recruit to the adult class), our stage-structured model predicted that
six of twelve translocated populations should experience a long-run
growth rate \(\lambda\) greater than 1 in the presence of Bd
(Figure~\ref{fig-viability}A; median predicted \(\lambda\) ranges from
1.15-1.36 for these six populations). These six populations all had
observed yearly adult survival greater than 0.5. When we reduced the
probability of successful recruitment \(\omega\) to 0.1 (a highly
conservative estimate), long-run growth rate of five of these six
populations was still greater than 1 (Figure~\ref{fig-viability}A).

Even when incorporating (i) demographic stochasticity and (ii)
environmental stochasticity in the probability of juvenile recruitment
to the adult class \(\omega\) (the transition that we expect to be the
most subject to environmental variability in the presence of Bd),
populations with high adult survival are likely to persist over a 50
year time horizon. Our model predicted that, following a single
introduction of 40 adult individuals into a population and using the
baseline value for \(\omega\) of 0.3, the six populations with the
highest adult survival probabilities (\(\sigma_{A_R} > 0.5\)) had
50-year extinction probabilities of 0 to 0.32
(Figure~\ref{fig-viability}B). This indicates strong potential for
long-term persistence in the presence of Bd and environmental
variability in recruitment. When we decreased \(\omega\) from 0.3 to a
conservative estimate of 0.15, 50-year extinction probability remained
less than 0.5 for three populations and was 0.79 to 0.94 for three
additional populations (Figure~\ref{fig-viability}B). In contrast, for
the six populations where yearly adult survival probability
\(\sigma_{A_R} < 0.5\), extinction probability over 50 years was always
predicted to be 1. To test the validity of our model predictions, we
demonstrated that our stochastic model could describe the general
recovery trajectory of our translocated population with the longest
survey history (Figure~\ref{fig-viability}C; population 70550, surveyed
for 16 years).

In summary, our model demonstrates that given observed yearly adult
survival probabilities of translocated frogs, 50\% of our translocated
populations have a high probability of population growth and long-term
viability in the presence of Bd. This is likely a conservative estimate
because there is evidence that naturally-recruited adults have higher
survival probability than translocated adults
(Figure S7), but we considered these
probabilities to be equal in all but three of our populations where we
had sufficient data to distinguish these different probabilities. In
addition, 5 of the 7 populations translocated after 2013 were predicted
to have long-term viability, compared to only 1 of 5 populations
translocated prior to 2013 (Table S1). Thus,
there is some trend suggesting that our ability to select high quality
habitats that promote long-term viability of translocated populations
improved over time.

\section*{Discussion}

Disease-induced population declines are decimating global biodiversity
\citep{daszak2000}, but strategies to recover affected species are
generally lacking. Here, we tested the possibility that populations of
resistant/tolerant individuals could be used to reestablish extirpated
populations. Although previously proposed
\citep{mendelson2019, scheele2021}, this strategy has rarely been used
in the wild \citep[see][ for a small-scale example]{joseph2018}. In
particular, although numerous studies have documented signatures of
selection in hosts following disease-induced declines
\citep[e.g.,][]{epstein2016, gignoux-wolfsohn2021}, we are unaware of
any that used individuals bearing the resulting adaptive alleles to
reestablish extirpated populations. The results of our study provide
evidence that (1) MYL frogs have experienced strong selection following
exposure to a recently-emerged lethal fungal disease and that some of
the genomic regions under selection contain genes related to disease
resistance/tolerance, and (2) reintroduction of individuals from these
resistant/tolerant populations allows the successful reestablishment of
viable frog populations despite ongoing disease.

Previous field studies in MYL frogs show that frog-Bd dynamics and frog
survival in the presence of Bd are fundamentally different between naive
and recovering populations. Following the arrival and establishment of
Bd in previously-naive populations, adult frogs develop high Bd loads
that lead to mass die-offs \citep{vredenburg2010}. In contrast, in
recovering populations adult frogs typically have low-to-moderate and
relatively constant Bd loads and mass die-offs are not observed
\citep[see also
Figure S5]{briggs2010, knapp2011}. The
differences in Bd load of frogs from naive and recovering populations
are also observed in controlled laboratory studies \citep[see Figure 4
in][]{knapp2016}. The observed reduced susceptibility of frogs from
recovering populations could in theory be due to several factors,
including natural selection for more resistant/tolerant genotypes and/or
acquired immunity, but until now evidence to evaluate the role of
evolution was lacking.

Results from our genomic analyses strongly suggest that natural
selection for adaptive alleles and resulting increased
resistance/tolerance is at least partially responsible for the reduced
susceptibility of frogs in recovering populations. We identified
specific alleles and genomic regions showing signatures of selection
between adjacent Bd-naive and recovering MYL frog populations,
consistent with strong selection following Bd exposure. Importantly,
some regions under selection are associated with cellular and
immunological mechanisms known to contribute to disease resistance,
including in amphibians \citep{zamudio2020}. For example, the MHC plays
an important role in immunity. In our study, we identified a region that
shows evidence of selection in recovering populations and contains eight
genes associated with either the MHC Class I or Class III regions. These
results corroborate numerous previous studies linking MHC genes to
amphibian resistance/tolerance against Bd
\citep[e.g.,][]{savage2011, bataille2015}. Similarly, the region with
the strongest indication of directional selection (as measured by
\(\pi_{diff}\)) contains the interferon-related gene GVINP1. Several
previous studies have found this gene to be differentially expressed
during Bd infection \citep[e.g.,][]{grogan2018, ellison2014} and in
populations differing in Bd susceptibility \citep{grogan2018}. This gene
is also strongly linked to disease in salmon, explaining a notable 20\%
of the resistance phenotype \citep{robledo2020, robledo2018}. We also
identified a region, characterized by high \emph{F\textsubscript{ST}}
and low \(\pi_{diff}\), that contained the complement genes C6 and C7.
The complement system plays an important role in innate immunity
\citep{rieraromo2016}, and our results could indicate that balancing
selection is acting in this region of the genome to favor a diverse set
of alleles, as is known for C6 in humans \citep{soejima2005}. Finally,
based on the analysis of individual outlier variants, the RIN3 gene
showed a consistent pattern of allele frequency differences across all
nine of the frog populations sampled in this study, indicating
consistent selection in populations distributed across a wide geographic
area. This gene is associated with inflammation response and in
\emph{Xenopus} is highly expressed in wound stumps following amputation
of tadpole tails \citep{fukazawa2009}. Also in this analysis, the
outlier variant with the lowest p-value (uncharacterized gene
LOC108802036) is located adjacent to a type I interferon gene (Np-IFNi2)
in another frog genome \citep{gan2018}, and together with GVINP1 further
suggests the importance of interferon-related genes in this system.

Reintroduction of resistant/tolerant MYL frogs was remarkably successful
in reestablishing viable populations in the presence of Bd. Of the 12
translocated populations, approximately 80\% showed evidence of both
successful reproduction and recruitment of new adults. Bd loads measured
before and after translocation were similar and generally remained well
below the level at which frog mortality occurs. Annual survival for 12
of the 24 translocated cohorts exceeded 50\%, and \textgreater{} 70\% of
translocated cohorts had survival above this 50\% level when the
earliest translocations are excluded (i.e., translocations conducted
when methods were still being refined). The fact that the relatively low
Bd loads and correspondingly high frog survival in the presence of Bd
was maintained between the natal/donor sites and the recipient sites
indicates that these characteristics of naturally-recovering populations
were not solely an effect of the natal/donor site or site * frog
interactions, but instead were due at least in part to
resistance/tolerance inherent in the frogs. In addition, the relatively
high survival of translocated frogs was maintained in their progeny, as
expected if resistance/tolerance has a genetic basis. Results from the
population viability model were also encouraging. In particular,
translocated populations with \textgreater{} 50\% survival in the first
year post-translocation were predicted to have a low probability of
extinction over 50 years (probability of extinction = 0 - 0.32). In
light of the generally low success rate of amphibian reintroduction
efforts \citep{dodd2005}, our success in reestablishing MYL frog
populations via translocation of resistant/tolerant individuals is
striking, and even more so given that this success occurred despite
ongoing infection of frogs with Bd.

Despite the resistance/tolerance of adult MYL frogs in recovering
populations, individual and population-level impacts of Bd are still
evident. In an earlier study of 2 of our 12 translocated populations
\citep{joseph2018}, Bd infection and load had detectable effects on the
survival of adults and may have influenced population establishment
(sites referred to as ``Alpine'' and ``Subalpine'' in \citep{joseph2018}
are identified as ``70550'' and ``70505'' in the current study).
Applying similar analyses to all 12 of our translocated populations
would likely provide a broader perspective of the ongoing effect of Bd.
In addition to these important but relatively subtle effects of Bd on
adults, the impacts on younger life stages are more apparent. MYL frogs
immediately following metamorphosis (``metamorphs'') are highly
susceptible to Bd infection \citep{ellison2018} and as a result
experience high mortality. This high susceptibility of metamorphs is
documented in numerous species of anurans, and may result from the
poorly developed immune system characteristic of this life stage
\citep{humphries2022}. In naturally recovering and translocated MYL frog
populations, we suggest that the high mortality of metamorphs is an
important limitation on subsequent recruitment of new adults. Therefore,
although adult MYL frogs appear relatively resistant/tolerant, Bd
infection continues to have important limiting effects on recovering
populations \citep[see also][]{hollanders2022}.

The recent emergence of Bd worldwide has contributed to the decline of
hundreds of amphibian species, some of which are now extinct in the wild
\citep{scheele2019}. This unprecedented impact on global amphibian
biodiversity is compounded by the lack of any effective and broadly
applicable strategies to reverse these impacts
\citep{garner2016, knapp2022}. Importantly, in addition to the natural
recovery documented for MYL frogs \citep{knapp2016}, other species are
also showing evidence of post-epizootic recovery in the presence of Bd
\citep{scheele2017, voyles2018} and suggest the possibility of also
using animals from these recovering populations to reestablish
extirpated populations. As with MYL frogs, the feasibility and long-term
success of such efforts will depend on the availability of robust donor
populations containing individuals that have acquired via evolution the
adaptive alleles necessary to allow frog survival and population growth
in the presence of Bd. Despite the hopeful example of successful
reestablishment of MYL frogs via evolutionary rescue and
reintroductions, the challenge of recovering hundreds of Bd-impacted
amphibian species worldwide is a daunting prospect. Although we now have
a proven strategy to reestablish extirpated MYL frog populations,
recovery across their large historical range will require substantial
resources over many decades. The results of this study provide a hopeful
starting point for that endeavor and other future efforts worldwide.

Evolutionary rescue may be increasingly critical for the survival of
wildlife populations in a rapidly changing world. The importance of
evolutionary rescue is known for a wide variety of taxa
\citep[e.g.,][]{carlson2014, mills2018, oziolor2019}, including in the
specific context of disease-impacted systems \citep{searle2020}.
However, obtaining evidence in wild animal populations for all the
causal links in this process (i.e., selection for adaptive alleles
\(\rightarrow\) increase in host fitness in the presence of the stressor
\(\rightarrow\) positive population growth and recovery) is difficult in
most systems. Nonetheless, we provide a compelling example from the wild
not only that evolutionary rescue can produce individuals that harbor
adaptive alleles and allow population recovery, but importantly that
these individuals can be used in translocations to reestablish
extirpated populations. We expect that such evolutionary rescue-based
translocations will be an essential tool in wildlife conservation in an
era of accelerating global change.

\matmethods{
\hypertarget{frog-evolution-in-response-to-bd-1}{%
\section*{Frog evolution in response to
Bd}\label{frog-evolution-in-response-to-bd-1}}

\hypertarget{sampling-and-sequencing}{%
\subsection*{Sampling and Sequencing}\label{sampling-and-sequencing}}

We collected DNA samples via buccal swabbing \citep{broquet2007} from 53
\emph{Rana muscosa}/\emph{Rana sierrae} individuals: 24 from 4 naive
populations, and 29 from 5 recovering populations. These populations are
located in the southern Sierra Nevada, from northern Yosemite National
Park to northern Sequoia National Park
(Figure~\ref{fig-allelefrequencies}). Samples were collected from 5-6
frogs per population. To minimize potential confounding effects caused
by variation in frog genotypes across latitude
\citep[e.g.,][]{rothstein2020}, we selected sampling sites such that
both population types were represented across similar latitudinal
ranges. DNA was extracted following Qiagen DNEasy manufacturer's
protocols. We sequenced the samples using an exome capture assay as
described in Byrne et al.~(in review). Briefly, genomic libraries were
prepared and captured using a custom Nimblegen capture pool. Capture
baits were designed based on the coding regions of the \emph{R. muscosa}
transcriptome (GenBank accession GKCT00000000). Captured libraries were
then pooled and sequenced on a NovaSeq 6000 150PE Flow Cell S1 at the
Vincent J. Coates Genomics Sequencing Lab at UC Berkeley. Raw sequencing
reads are available from NCBI SRA (PRJNA870451).

\hypertarget{data-pre-processing-and-cleaning}{%
\subsection*{Data Pre-processing and
Cleaning}\label{data-pre-processing-and-cleaning}}

Raw reads were filtered for adapters and contaminants using fastp
\citep{chen2018} and aligned to the \emph{Rana muscosa} genome (NCBI
SRA: SRS6533475 \citep{hon2020}), with repetitive elements masked using
bwa (``mem'' mode) \citep{li2013}. Exact PCR duplicates were marked
using Picard. Variants were then called following GATK best practices
(v.4.2.0.0 \citep{van2020}). Briefly, raw variants were called for each
sample using HaplotypeCaller and combined using CombineGVCFs. Next,
genotypes were jointly called using GenotypeGVCFs. Variants were then
hard filtered using gatk VariantFiltration using the following
parameters to remove low-quality sites: QD \textless{} 2.0, FS
\textgreater{} 40.0, SOR \textgreater{} 3.0, MQ \textless{} 50.0,
MQRankSum \textgreater{} 3.0, MQRankSum \textless{} -3.0, ReadPosRankSum
\textgreater{} 3.0, ReadPosRankSum \textless{} -3.0. This initial filter
resulted in 1,595,206 variant sites across 53 individuals.

We then further filtered our dataset at the individual and variant
level. First, we trimmed our variants to only include those with minor
allele frequency \textgreater{} 0.03, a maximum depth of 250 and minimum
depth of 5, a minimum genotype quality of 20, and a maximum missing
proportion of 0.5. This filter resulted in 427,038 sites, of which
353,172 were SNPs and 73,866 were INDELS. Finally, we trimmed samples
with an average depth across filtered sites \textless{} 7x. Our final
dataset included 50 samples, 23 from naive and 27 from recovering
populations, with an average depth of 16.7x (range = 7.4x -- 26.1x).

\hypertarget{data-analysis}{%
\subsection*{Data Analysis}\label{data-analysis}}

To visualize the genomic relationships of our populations we conducted a
PCA using the glPCA function in the adegenet R package
\citep{jombart2008}. To detect regions of the genome under selection
between naive and recovering populations, we used two approaches: (1) a
multivariate linear mixed model to evaluate individual variants (SNPs
and INDELs), and (2) a splined window analysis to evaluate larger
genomic regions. For the variant analysis, we first used a stringent
data filter to include only variants with \textless{} 5\% missing data
(missing for no more than 2 individuals), and then calculated the
likelihood ratio statistic for the resulting set of 148,307 high quality
variants across 127 contigs using GEMMA \citep{zhou2014}. GEMMA
calculates and incorporates a relatedness matrix for input samples,
allowing us to account for relatedness and population structure when
calculating likelihood ratio statistics. We identified variants showing
different allele frequencies between naive versus recovering populations
(``outliers'') using a Bonferroni-corrected significance level of 0.01.
We visualized the results using a Manhattan plot and qqplot. We
developed a more liberal set of outlier variants using a
Bonferroni-corrected significance level of 0.05 and used this set solely
for the gene ontology (GO) analysis (see below and SI Appendix: Dataset
S1, S2).

In the analysis of individual variants, for each outlier variant we
determined whether the variant was synonymous (protein sequence the same
for each variant) or non-synonymous (protein sequence differs between
variants), and where in the gene it was located. To do this, we first
extracted the reference genome sequence surrounding the variant using
the bedtools ``getfasta'' function \citep{quinlan2010}. Next, we
re-annotated each sequence using BLAST to get the predicted gene
location based on the closest annotated reference \citep{altschul1997}.
We then translated each variant to amino acids and aligned this
translation to that of the gene annotation to ensure proper frame of
reference using Geneious \citep{kearse2012}. After ensuring proper
translation, we characterized variants as within or outside the coding
sequence of the gene and as either synonymous or non-synonymous.

In the splined window analysis, we identified outlier regions using
\emph{F\textsubscript{ST}} and differences in nucleotide diversity
(\(\pi_{diff}\)) between naive and recovering populations. First, we
calculated per-site \emph{F\textsubscript{ST}} between the naive and
recovering individuals for all bi-allelic SNPs in the 30 largest contigs
(98\% of all SNPs) using VCFtools \citep{danecek2011}. Next, we
calculated per-site nucleotide diversity \(\pi\) separately for
individuals from the naive and recovering populations using VCFtools,
then calculated \(\pi_{diff}\) for each population
(\(\pi_{diff} = \pi_{naive} - \pi_{recovering}\)). We concatenated the
values for \emph{F\textsubscript{ST}} and \(\pi_{diff}\) in order of
size-sorted chromosome number and adjusted the SNP position based on the
relative position in the genome. We then used the GenWin R package
\citep{beissinger2015} to conduct a splined discrete window analysis for
\emph{F\textsubscript{ST}} and \(\pi_{diff}\). This method calculates
where non-overlapping window boundaries should occur by identifying
inflection points in the fitted spline, therefore balancing false
positive and false negative results that occur using other window-based
methods \citep{beissinger2015}. This method also calculates a
W-statistic allowing for outlier identification. We identified outliers
as those with a W-statistic greater than 4 standard deviations above the
mean for \emph{F\textsubscript{ST}} or above/below the mean for
\(\pi_{diff}\). Shared outliers were then identified as those that were
outliers in both analyses, meaning that they showed (1) high
differentiation between naive and recovering populations, and (2)
differential patterns of nucleotide diversity in the same region.
Finally, we extracted gene transcripts mapped within each region and
retrieved annotation for that region using BLAST (SI Appendix: Datasets
S3, S4).

Using the liberal set of outlier variants, we performed a statistical
overrepresentation test (Fisher's exact test) using Blast2GO
\citep{conesa2005}. First, we retrieved the BLAST hits and mapped GO
terms for each gene in our targeted transcriptome. Next, we compared the
30 unique outlier genes from the liberal set of outlier variants to the
complete set of genes in our target transcriptome to determine if any GO
biological functions, molecular functions, or cellular processes were
overrepresented. We then repeated this process for the set of 33 genes
located in the overlapping \emph{F\textsubscript{ST}} and \(\pi_{diff}\)
splined windows. Code used for genomic analyses and to create figures is
available from the following GitHub repository:
\url{https://github.com/allie128/mylf-selection}.

\hypertarget{frog-population-recovery-1}{%
\section*{Frog population recovery}\label{frog-population-recovery-1}}

\hypertarget{field-methods}{%
\subsection*{Field methods}\label{field-methods}}

For the 24 translocations we conducted, we identified donor populations
from which adult frogs (\(\geq\) 40 mm snout-vent length) could be
collected using several years of visual encounter surveys (VES) and skin
swab collections \citep{knapp2016}, and results from population genetic
analyses \citep{poorten2017}. The populations that we selected contained
several hundred \emph{R. sierrae} adults and hundreds or thousands of
tadpoles. These relatively high abundances were the result of recent
increases following previous Bd-caused declines \citep{knapp2016}. As is
typical for recovering MYL frog populations, Bd prevalence in the donor
populations was high (0.69--0.96) and Bd load (median
log\textsubscript{10}(load) = 3.06--3.78 ITS copies) was two or more
orders of magnitude below the level at which frog mortality is expected
(log\textsubscript{10}(load) \(\approx\) 5.78
copies)\citep{vredenburg2010, joseph2018}. Recipient sites to which
frogs were translocated were chosen based on previous \emph{R. sierrae}
presence (determined from VES and/or museum records) or characteristics
that suggested high quality habitat for this species \citep{knapp2005}.

We conducted 1--4 translocations per site (Figure S3,
Figure~\ref{fig-translocation-survival}) and each translocated cohort
included 18 to 99 frogs (median = 30). In preparation for each
translocation, adult frogs were collected from the donor population and
measured, weighed, swabbed, and PIT tagged. Frogs were transported to
the recipient site either on foot or via helicopter. Following release,
translocated populations were visited approximately once per month
during the summer active season (median number of visits per summer = 2;
range = 1--10) and assessed using diurnal CMR surveys and VES. CMR
surveys allowed estimation of adult survival, adult population size, and
recruitment of new adults, and VES provided estimates of tadpole and
juvenile abundance. During 2006-2012, CMR surveys were conducted on a
single day (primary period) per site visit, during which all habitats
were searched repeatedly for adult frogs. Frogs were captured using
handheld nets, identified via their PIT tag (or tagged if they were
untagged), measured, weighed, swabbed, and released at the capture
location. During 2013-2022, we generally used a robust design in which
all habitats were searched during several consecutive days (median
number of secondary periods per primary period = 3; range = 3--7), and
frogs were processed as described above. However, when the number of
frogs detected on the first survey day was zero or near zero, we
typically conducted only a single-day CMR survey. When using a robust
design, within a primary period, frogs that were captured during more
than one secondary period were measured, weighed, and swabbed during the
first capture, and during subsequent captures were only identified and
released.

During each site visit, we conducted VES either immediately before CMR
surveys or during the first day of CMR surveys. VES was conducted by
walking the entire water body perimeter, first 100 m of each inlet and
outlet stream, and any fringing ponds and wetlands, and counting all
\emph{R. sierrae} tadpoles and juveniles. These \emph{R. sierrae} life
stages have high detectability, and counts are highly repeatable and
provide estimates of relative abundance \citep{knapp2000}.

\hypertarget{frog-counts-and-reproductive-success}{%
\subsection*{Frog counts and reproductive
success}\label{frog-counts-and-reproductive-success}}

For each of the translocated populations, we used the presence of
tadpoles and/or juveniles from VES and counts of new recruits (i.e,
untagged adults) in CMR surveys to provide two measures of successful
reproduction. To calculate the proportion of years in which
tadpoles/juveniles were present at a site, we excluded surveys conducted
in the year of the initial translocation to that site. This exclusion
accounted for the fact that all translocations were conducted after the
breeding period and reproduction would therefore not have occurred until
the following year. Similarly, to calculcate the proportion of years in
which new recruits were present at a site, we excluded surveys conducted
during the 3 years following the initial translocation. This accounted
for the multi-year tadpole and juvenile stages in MYL frogs
(Table S2).

\hypertarget{estimation-of-frog-survival-and-abundance}{%
\subsection*{Estimation of frog survival and
abundance}\label{estimation-of-frog-survival-and-abundance}}

We estimated survival of translocated frogs, recruitment of new frogs
into the adult population, and adult population size using a Bayesian
open-population Jolly-Seber mark-recapture model with known additions to
the population (i.e., translocated cohorts), as implemented by the mrmr
package \citep{joseph2019} and using R Statistical Software
\citep[v4.4.4,][]{rsoftware2022}. The model tracks the states of
\emph{M} individuals that comprise a superpopulation made up of real and
pseudo-individuals \citep[see][ for details]{joseph2018}. The possible
states of individuals include ``not recruited'', ``alive'', and
``dead''. The possible observations of individuals include ``detected''
and ``not detected''. We assume that individuals that are in the ``not
recruited'' or ``dead'' states are never detected (i.e., there are no
mistakes in the individual PIT tag records). For all models, we used
mrmr defaults for priors, number of chains (4), and warmup and
post-warmup iterations (2000 for each). We evaluated convergence of the
Markov chain Monte Carlo (MCMC) algorithm using trace plots and
Gelman-Rubin statistics (Rhat).

\hypertarget{predictors-of-post-translocation-frog-survival}{%
\subsection*{Predictors of post-translocation frog
survival}\label{predictors-of-post-translocation-frog-survival}}

To identify important predictors of frog survival following
translocation, we developed linear and multilevel models in a Bayesian
framework. We performed all analyses with the rstanarm package
\citep{rstanarm2022} and R Statistical Software
\citep[v4.4.4,][]{rsoftware2022}, and used default, weakly informative
priors. Starting models included all relevant population-level effects
and, when data contained possible hierarchical structure, we included
group-level effects in the full model. We evaluated model fit using
leave-one-out (``LOO'') cross-validation
\citep{gelman2013, gabry2019, vehtari2016}, as implemented by the loo
package \citep{vehtari2022}.

We conducted two analyses. In the first, we addressed the more general
question of whether survival of frogs in the first translocation to a
site predicted the survival of frogs in subsequent translocations.
Group-level effects included site\_id or translocation\_id. The second
analysis focused specifically on identifying predictors of
post-translocation frog survival using a wide range of variables
describing characteristics of frogs, sites, and translocations
(predictor variables: Bd load, sex, frog size, site elevation, winter
severity in the year of translocation, winter severity in the year
following translocation, donor population, day of year on which a
translocation was conducted, and translocation order). In both analyses
we used 1-year survival as a single metric of survival that was
available for all translocated populations regardless of how many years
of post-translocation data were available (range = 3--16). Group-level
effects included site\_id, translocation\_id, or translocation\_id
nested within site\_id. In both analyses, the response variable,
probability of 1-year frog survival, was rounded to integer values to
produce a binomial outcome, and all models used a bernoulli family. For
all models, we used default priors, four chains, and 5000 iterations
each for warmup and post-warmup. We checked convergence of the MCMC
algorithm using trace plots and Rhat.

\hypertarget{changes-in-bd-load-following-translocation}{%
\subsection*{Changes in Bd load following
translocation}\label{changes-in-bd-load-following-translocation}}

We analyzed skin swabs using standard Bd DNA extraction and qPCR methods
\citep[see SI Appendix for details]{boyle2004}. To assess the magnitude
of changes in Bd load on frogs following translocation, we compared Bd
loads measured before versus after translocation. Before-translocation
loads were quantified using skin swabs collected from all
to-be-translocated frogs at the donor site on the day before or the day
of the translocation. After-translocation Bd loads were based on all
swabs collected from translocated frogs at the recipient site in the
year of and the year following translocation. Individual frogs and their
associated Bd loads were included in the dataset only if frogs were
captured at the recipient site at least once during the year 0-1 period.
We used an analysis workflow similar to that described above. Briefly,
the response variable was before-translocation load and the predictor
variable was after-translocation load. We compared models with and
without group-level effects, and group-level effects included
pit\_tag\_id, pit\_tag\_id nested within site\_id, or pit\_tag\_id
nested within translocation\_id. All models used a negative binomial
family. Priors, chains, iterations, and checks were as described above.

Code to replicate all of the analyses described above is available from
the following GitHub repositories:
\url{https://github.com/SNARL1/translocation};
\url{https://github.com/SNARL1/cmr-analysis}.

\hypertarget{population-viability-modeling}{%
\section*{Population viability
modeling}\label{population-viability-modeling}}

\hypertarget{model-description}{%
\subsection*{Model description}\label{model-description}}

To determine the implications of observed 1-year adult survival on the
long-term viability of populations established via translocation, we
developed a population model for MYL frogs. Our central question was:
How does the magnitude and variation in observed adult survival
probability across translocated populations affect the long-term
persistence probability of populations? We developed a model that
tracked seven state variables of a frog population: density of
translocated adults (\(A_T\)), density of adults naturally recruited
into the population (\(A_R\)), density of first-year tadpoles (\(L_1\)),
density of second-year tadpoles (\(L_2\)), density of third-year
tadpoles (\(L_3\)), density of first-year juveniles (\(J_1\)), and
density of second-year juveniles (\(J_2\)). We divided adults into two
classes \(A_T\) and \(A_R\) because there is evidence that the survival
probability of translocated adults and naturally recruited adults
differs (Figure S7).

We modeled the dynamics of these seven state variables using a
discrete-time, stage-structured model where a time step is one year. The
dynamics are given by equation \ref{eqn:matrix}.

\begin{figure*}\[
\begin{bmatrix}
L_1 \\
L_2 \\
L_3 \\
J_1 \\
J_2 \\
A_R \\ 
A_T
\end{bmatrix}(t + 1) = 
\begin{bmatrix}
  0 & 0 & 0 & 0 & 0 & \sigma_{A_R} p_F F & \sigma_{A_T} p_F F \\
  \sigma_{L1} p_{L1} & 0 & 0 & 0 & 0 & 0 & 0 \\
  0 & \sigma_{L2} p_{L2} & 0 & 0 & 0 & 0 & 0\\
  \sigma_{L1} (1 - p_{L1}) & \sigma_{L2} (1 - p_{L2}) & \sigma_{L3} & 0 & 0 & 0 & 0 \\
  0 & 0 & 0 & \sigma_{J1} p_{J1} & 0 & 0 & 0 \\
  0 & 0 & 0 & \sigma_{J1} (1 - p_{J1}) \omega & \sigma_{J2} \omega & \sigma_{A_R} & 0 \\
  0 & 0 & 0 & 0 & 0 & 0 & \sigma_{A_T} 
\end{bmatrix} \begin{bmatrix}
L_1 \\
L_2 \\
L_3 \\
J_1 \\
J_2 \\
A_R \\
A_T \\
\end{bmatrix}(t)\numberthis \label{eqn:matrix} 
\]\end{figure*}

The parameters in this model are yearly survival probability
\(\sigma_{\cdot}\) (the subscript ``\(\cdot\)'' indicates a particular
state variable), probability that a female frog reproduces in a given
year \(p_F\), number of eggs produced by a female frog in a year that
successfully hatch \(F\), probability of a first-year tadpole remaining
as a tadpole \(p_{L1}\), probability of a second-year tadpole remaining
as a tadpole \(p_{L2}\), and probability of a first-year juvenile
remaining as a juvenile \(p_{J1}\), and the probability of a juvenile
successfully recruiting to an adult \(\omega\).

In this model we ignore density-dependent recruitment because we were
interested in the growth of the population from an initial
reintroduction and whether this growth was sufficient to prevent
extinction over 50 years following the introduction. We also did not
directly consider the dynamics of Bd in this model. We made this
decision because (i) translocated populations are infected with Bd at
high prevalence \citep{joseph2018}, and (ii) host density does not seem
to play a significant role in multi-year Bd infection dynamics in this
system \citep{wilber2022}. Thus, ignoring Bd infection dynamics and
instead assuming all host vital rates are in the presence of high Bd
prevalence significantly simplifies the model without much loss of
realism. Additional details are provided in the SI Appendix.

\hypertarget{model-analysis}{%
\subsection*{Model analysis}\label{model-analysis}}

After parameterizing our model with CMR-estimated adult frog survival
probabilities and other known vital rates
(Table S2), we performed four analyses. First, we
compute the long-run growth rate \(\lambda\) for each of our 12
translocated populations to determine if the populations were
deterministically predicted to grow or decline in the long-run. Second,
we compute the elasticity of \(\lambda\) to model parameters, to
quantify how much changes in adult survival, for example, affected the
long-run growth rate. Third, we included demographic and environmental
stochasticity in our model and simulated the 50-year viability (i.e., 1
- extinction probability) of populations given an introduction of 40
adult individuals into an unoccupied habitat. Finally, we fit our model
to our longest translocation trajectory to confirm that our model could
reasonably reproduce the observed recovery trajectories of MYL frogs
following reintroductions. Additional details are provided in the SI
Appendix. Code to replicate the analyses can be found at
\url{https://github.com/SNARL1/translocation}.

}

\showmatmethods{} % Display the Materials and Methods section

\acknow{We thank the following for important contributions to this study: E.
Hegeman and A. Lindauer--project and data management, field and
laboratory work; A. Barbella and K. Rose--laboratory work; C. Kamoroff
and numerous summer technicians--field work; staff at Sequoia-Kings
Canyon and Yosemite National Parks, Inyo and Sierra National Forests,
California Department of Fish and Wildlife, U.S. Fish and Wildlife
Service, University of California-Santa Barbara Institutional Animal
Care and Use Committee, and Sierra Nevada Aquatic Research
Laboratory--research permits, logistical support, and/or field
assistance. This project was supported by grants from the National Park
Service (to R.A.K.), Yosemite Conservancy (to R.A.K.), and National
Science Foundation (EF-0723563, to C. Briggs; DEB-1557190, to C. Briggs;
DEB-2133401, to M.Q.W; and DBI-2120084, to C. Richards-Zawacki).}

\showacknow{} % Display the acknowledgments section

% Bibliography
\bibliography{translocation}

\clearpage
\begin{figure*}

{\centering \includegraphics[width=\textwidth]{figures/allele_maps.png}

}

\caption{\label{fig-allelefrequencies}Evidence for selection on
individual variants in recovering MYL frog populations at the landscape
scale. For each of the 9 naive and recovering MYL frog populations
(indicated by numbered points), adjacent pie charts show allele
frequencies for the 11 outlier SNPs from 7 distinct genes: (A)
LOC108802036, (B) TCF19, VARS, MAP1S, TYRP1, and SRFBP1, and (C) RIN3.
Charts are superimposed on a map of the Sierra Nevada study area, with
Yosemite, Kings Canyon, and Sequoia National Parks (from north to south)
shown in dark gray, and the range boundary for MYL frogs shown in light
gray. The inset map locates the study area and range boundary in
California.}

\end{figure*}

\newpage

\begin{figure*}

{\centering \includegraphics[width=\textwidth]{figures/splinewindow_manhattan.png}

}

\caption{\label{fig-spline-manhattan}Evidence for selection on genomic
regions in recovering MYL frog populations. Manhattan plot of the
results from the splined window analysis showing outlier regions for the
difference in (A) nucleotide diversity \(\pi_{diff}\) and (B)
\emph{F\textsubscript{ST}}. In (A), outlier regions are shown above the
upper red dashed line and below the lower red dashed line. In (B),
outlier regions are shown above the single dashed red line. Outlier
regions for either \(\pi_{diff}\) or \emph{F\textsubscript{ST}} are
shown in blue and outlier regions for both \(\pi_{diff}\) and
\emph{F\textsubscript{ST}} are shown in red. (C) Magnified Contig19 from
(A) showing two adjacent outlier regions for \(\pi_{diff}\) 12.9Mb
upstream of the RIN3 outlier SNP (indicated with a dashed vertical blue
line). (D) Magnified Contig8 from from (B) showing the
\emph{F\textsubscript{ST}} outlier region that includes the outlier SNPs
TCF19 and VARS. This region of the genome contains 8 annotated genes
known to occur in the extended MHC Class I and III regions.}

\end{figure*}

\newpage

\begin{figure*}

{\centering \includegraphics[width=0.5\textwidth]{figures/translocation_survival_bysiteid.png}

}

\caption{\label{fig-translocation-survival}Evidence for frog survival at
the recipient sites following translocation. Median 1-year survival for
each cohort of translocated frogs at the 12 recipient sites. Error bars
show the 95\% uncertainty intervals. Sites are arranged along the x-axis
using the average of the median 1-year survival per translocation at
each site. Dot colors indicate the donor population from which frogs in
each translocated cohort were collected. When multiple translocations
were conducted to a site, points and error bars are slightly offset to
avoid overlap.}

\end{figure*}

\newpage

\begin{figure*}

{\centering \includegraphics[width=\textwidth]{figures/cond_effects_plot.png}

}

\caption{\label{fig-cond-effects}Evidence for reduced impacts of Bd on
MYL frogs as indicated by the lack of an effect of Bd load on survival.
Results show conditional effects of the important predictors of 1-year
frog survival (survival is expressed as a probability). (A) winter
severity in the year following translocation, (B) elevation of recipient
site, (C) donor population, and (D) sex. In (A) and (B), blue lines are
medians and gray ribbons are 95\% uncertainty intervals. In (C) and (D),
box plots show medians, first and third quartiles, largest and smallest
values within 1.5x interquartile range, and values outside the 1.5x
interquartile range.}

\end{figure*}

\newpage

\begin{figure*}

{\centering \includegraphics[width=\textwidth]{figures/pop_viability_figures_for_manuscript.jpg}

}

\caption{\label{fig-viability}Population viability analysis of
translocated MYL frogs quantified the long-run growth rate and
extinction probability of frog populations: (A) Predicted long-run
growth rate \(\lambda\) for different values of yearly adult survival
probability \(\sigma_{A_R}\) and probability of recruitment \(\omega\),
given the parameterized, deterministic model. Colored points show the
predicted \(\lambda\) values for the twelve translocated populations
when probability of recruitment into the adult class is \(\omega = 0.3\)
(indicated by the dashed line). The red line shows where
\(\lambda = 1\). Note that the point for 70413 is mostly hidden behind
other points. (B) Predicted 50-year extinction probabilities of the 12
translocated populations, given demographic stochasticity, environmental
variability in \(\omega\), and different mean values of \(\omega\). The
dashed line shows where \(\omega = 0.3\), for comparison to (A). There
are 6 lines at extinction probability = 1, 5 of which (70414--70619) are
hidden beneath the line for 70505. (C) 100 simulated trajectories (black
lines) from the population viability model that most closely matched the
observed abundance trajectory of adult amphibians at site 70550 (light
orange).}

\end{figure*}

\newpage
\end{document}