### 2024-04-21 doc/manuscript/translocation_bioarxiv_SI.tex

\documentclass[9pt,twoside,lineno]{pnas-new-SI}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnassupportinginfo}
\usepackage{longtable}
\usepackage{blkarray}

\title{Reintroduction of resistant frogs facilitates landscape-scale recovery in the presence of a lethal fungal disease}
\author{Roland A. Knapp, Mark Q. Wilber, Allison Q. Byrne, Maxwell B. Joseph, Thomas C. Smith, Andrew P. Rothstein, Robert L. Grasso, Erica Bree Rosenblum}
\correspondingauthor{Corresponding Author name. Roland A. Knapp \\ E-mail: roland.knapp@ucsb.edu}

\begin{document}

\maketitle

%% Adds the main heading for the text. Comment out this line if you do not have any supporting information text.
\SItext
\hypertarget{frog-population-recovery-2}{%
\section{Frog population
recovery}\label{frog-population-recovery-2}}

\hypertarget{laboratory-methods}{%
\subsection{Laboratory methods}\label{laboratory-methods}}

Swab extracts were analyzed using standard Bd DNA extraction and qPCR
methods \citep{boyle2004}, and extracts were analyzed singly instead of
in triplicate \citep{kriger2006}. For analysis of swabs collected during
2005--2014, we used standards developed from known concentrations of
zoospores \citep{boyle2004}, and after 2014, we used standards based on
single ITS1 PCR amplicons \citep{longo2013}. Based on paired comparisons
between samples analyzed using both types of standards, Bd in the study
area has an average of 60 ITS1 copies per zoospore. To express all qPCR
results as the number of ITS1 copies, starting quantities obtained using
the zoospore standard (measured as ``zoospore equivalents'') were
multiplied by 60. In addition, all qPCR quantities (regardless of
standard) were multiplied by 80 to account for the fact that DNA
extracts from swabs were diluted 80-fold during extraction and PCR
\citep{vredenburg2010}.

\hypertarget{cmr-model-structure}{%
\subsection{CMR model structure}\label{cmr-model-structure}}

We estimated survival and recruitment for each site using open
population CMR models based on \citep{joseph2018}. For each individual
in a superpopulation of real and psuedo-individuals, \(i=1, ..., M\) on
each survey \(j=1, ..., n_j\): \(o_{i, j} = 1\) if the individual was
not detected, and \(o_{i, j}=2\) if the individual was detected. Capture
histories of \(M\) individuals are modeled, although only \(N_s\)
individuals were captured. This parameter expanded data augmentation
allows us to account for the possibility that undetected individuals may
have recruited into the adult population \citep{royle2012}. Here, \(M\)
was chosen to be three times the number of observed individuals
(\(3N_s\)) to be considerably greater than our prior guess of \(N_s\).

We denote the true state of individual \(i\) as \(u_{i, t}\) for primary
period \(t = 1,..., n_t\). The states that we consider are:
\(u_{i, t} = 1\) for individuals that have not recruited,
\(u_{i, t} =2\) for live adults, and \(u_{i, t} = 3\) for dead adults.
Each survey \(j=1, ..., n_j\) occurs in one of the \(n_t\) primary
periods, and we denote the primary period in which survey \(j\) takes
place as \(t_j\). Multiple primary periods can occur during a summer
active season (typically July-August, but in some years as early as May
and as late as September). We set the year containing the first primary
period to \(y_{t = 1} = 1\), and generally \(y_t\) represents the year
containing primary period \(t\). Years increment by one until the final
year of the mark recapture efforts, which we denote \(n_y\):
\(y \in \{1, 2, ..., n_y\}\). We assume that within a primary period,
the state of each individual does not change (i.e., individuals do not
recruit into the adult population, gain or lose Bd infection, or die).
This assumption is justified by the short time intervals between surveys
within primary periods, in cases where primary periods contain multiple
secondary periods.

Live individuals are detected with probability \(p_j\), which is modeled
as:

\[p_j = \text{logit}^{-1}(X_j^{(p)} \beta^{(p)}),\]

where \(X_j^{(p)}\) is a known row vector and \(\beta^{(p)}\) an unknown
parameter vector. Not recruited and dead individuals are never captured.
We bundle these assumptions about the observation probabilities for
survey \(j\) into an emission matrix \(\Omega_j\):

\[
\Omega_j =
\begin{blockarray}{ccc}
  \text{Not detected} & \text{Detected} \\
\begin{block}{(cc)c}
  1 & 0 & \text{Not recruited} \\
  1 - p_j & p_j & \text{Alive} \\
  1 & 0 & \text{Dead} \\
\end{block}
\end{blockarray}
\]

The state transition matrix \(\Psi_{t, i}\) contains the probabilities
of individual \(i\) transitioning from state \(u_{i, t}\) (rows) to
\(u_{i, t+1}\) (columns) between primary period \(t\) and \(t+1\). For
non-introduced (i.e., naturally recruited) individuals, this matrix is
given by:

\[
\Psi_{t, i} =
\begin{blockarray}{cccc}
  \text{Not recruited} & \text{Alive} & \text{Dead} \\
\begin{block}{(ccc)c}
  1 - \lambda_t & \lambda_t & 0 & \text{Not recruited} \\
  0 & \phi_t & 1 - \phi_t & \text{Alive} \\
  0 & 0 & 1 & \text{Dead} \\
\end{block}
\end{blockarray},
\] where \(\lambda_t\) is the probability of recruiting in time \(t\)
and \(\phi_t\) is the probability of survival in time \(t\).

For introduced individuals, which have deterministic recruitment (i.e.,
they recruit when introduced), the state transition matrix is given by:

\[
\Psi_{t, i} =
\begin{blockarray}{cccc}
  \text{Not recruited} & \text{Alive} & \text{Dead} \\
\begin{block}{(ccc)c}
  1 - I_{t, i} & I_{t, i} & 0 & \text{Not recruited} \\
  0 & \phi_t & 1 - \phi_t & \text{Alive} \\
  0 & 0 & 1 & \text{Dead} \\
\end{block}
\end{blockarray},
\] where \(I_{t, i}\) is a known indicator function for whether
individual \(i\) was introduced in primary period \(t\).

We allow recruitment probabilities to vary in time via random effects,
such that:

\[\lambda_t = \text{logit}^{-1}(\alpha^{(\lambda)} + \epsilon^{(\lambda)}_t),\]

where \(\alpha^{(\lambda)}\) is an intercept parameter and
\(\epsilon^{(\lambda)}_t\) is an adjustment for time \(t\).

Survival probabilities also vary in time, and as a function of known
covariates:

\[\phi_t = \text{logit}^{-1}(X^{(\phi)}_t \beta^{(\phi)} + \epsilon^{(\phi)}_t),\]

where \(X^{(\phi)}_t\) is a row vector of known covariates,
\(\beta^{(\phi)}\) is a column vector of unknown coefficients, and
\(\epsilon^{(\phi)}_t\) is an adjustment for time \(t\).

To complete the specification of the Bayesian model, we specify priors
for all unknown parameters. The recruitment parameter priors were
specified as follows:

\[\alpha^{(\lambda)} \sim N(0,1),\]
\[\sigma^{(\lambda)} \sim N_+(0,1),\]
\[\epsilon^{(\lambda)}_t \sim N(0, \sigma^{(\lambda)}),\] for periods
\(t=1, ..., T\). Here \(N\) represents the normal distribution and
\(N_+\) the half normal distribution with positive support.

Survival parameter priors were specified similarly as:

\[\beta^{(\phi)} \sim N(0, 1),\] \[\sigma^{(\phi)} \sim N_+(0, 1),\]
\[\epsilon^{(\phi)}_t \sim N(0, \sigma^{(\phi)}),\] for time
\(t=1, ..., T\).

The detection model coefficient vector also received a standard normal
prior \(\beta^{(p)} \sim N(0, 1)\).

We computed the likelihood of each individual capture history using the
forward algorithm, and we estimated the latent states using the
forward-backward algorithm \citep{zucchini2009, joseph2018}.

All of the code to specify and fit the model in Stan is available in the
open source mrmr package \citep{joseph2019}.

The joint distribution of the resulting model can be written as follows:

\begin{multline*}
[\alpha^{(\lambda)}, \sigma^{(\lambda)}, \epsilon^{(\lambda)}_{1:T}, \beta^{(\phi)}, \sigma^{(\phi)}, \epsilon^{(\phi)}_{1:T}, \beta^{(p)} \mid \pmb Y] \propto \prod_{i=1}^M [Y_i \mid \alpha^{(\lambda)}, \epsilon^{(\lambda)}_{1:T}, \beta^{(\phi)}, \epsilon^{(\phi)}_{1:T}, \beta^{(p)}] \times \\
\prod_{t=1}^T [\epsilon_t^{(\lambda)} \mid \sigma^{(\lambda)}] [\epsilon_t^{(\phi])} \mid \sigma^{(\phi)}] [\sigma^{(\lambda)}] [\sigma^{(\phi)}] [\alpha^{(\lambda)}] [\beta^{(\phi)}] [\beta^{(p)}],
\end{multline*}

where \(\pmb Y\) is an \(M \times T\) detection matrix, and \(Y_i\) the
capture history of individual \(i\).

\hypertarget{among-site-survival-modeling}{%
\subsection{Among-site survival
modeling}\label{among-site-survival-modeling}}

The objective of this analysis is to describe the influence of site,
cohort, and individual level characteristics on post-translocation frog
survival. By modeling survival estimates obtained from site-specific
mrmr CMR analyses, we are in effect conducting an among-site
meta-analysis. Although it would theoretically be possible to estimate
survival covariate effects in a joint CMR model that integrates capture
histories across all sites, this was impractical due to computational
requirements of the CMR models (namely, run time and memory).

We used Bayesian generalized linear mixed models to investigate
predictors of survival among sites. The response \(y_i\) is binary,
representing a point estimate of whether individual \(i\) survived in
the year following translocation. We generated these point estimates by
rounding the posterior median of 1-year post-introduction survival for
each individual (from site-specific mrmr CMR models) and modeled the
data using a Bernoulli distribution:

\[y_i \sim \text{Bernoulli}(p_i),\]

where \(p_i\) is the probability of survival.

We modeled variation in probabilities as follows:

\[\text{logit}(p_i) = \alpha + X_i \pmb \beta + \pmb \nu_{g[1:N]},\]

where \(\alpha\) is an intercept, \(X_i\) is a length \(K\) row vector
of predictors, \(\pmb \beta\) is a column vector of predictor effects,
and \(\pmb \nu\) a vector of group level random effects. Here \(g[i]\)
refers to the group \(g\) containing individual \(i\), and we estimate
an adjustment for each of the \(G\) groups (\(\nu_1, ..., \nu_G\)).

These models were fit using the \texttt{stan\_glmer()} function in the
rstanarm package, with default priors described below
\citep{rstanarm2022}. These priors are vague, but include data-dependent
scaling as follows to account for different input variable scales.
However, because we standardized all predictor variables similarly to
have equal variance (by centering and dividing by twice the sample
standard deviation), the resulting priors are identical. Specifically,
we have:

\[\alpha \sim \text{Normal}(0, 2.5),\] and

\[\beta_k \sim \text{Normal}(0, 5),\]

for \(k=1, ..., K\) where \(K\) is the number of predictor variables.

The default prior for group level adjustments \(\nu_1, ..., \nu_G\) in
rstanarm is a zero-mean Gaussian, where the covariance matrix is
constructed from a correlation matrix with an LKJ prior, and a vector of
variance parameters -- the decomposition of variance prior with unit
regularization, concentration, shape, and scale parameters
\citep{lewandowski2009}.

We drew posterior samples using Dynamic Hamiltonian Monte Carlo in Stan,
with four parallel chains, each run for 10,000 iterations, discarding
the first half of each chain as warm-up draws \citep{rstanarm2022}. We
used Rhat statistics and trace plots to verify convergence. We
considered models with different subsets of fixed and random effects,
and used approximate leave-one-out cross validation to identify the best
model \citep{vehtari2016}.

\hypertarget{population-viability-modeling-1}{%
\section{Population viability
modeling}\label{population-viability-modeling-1}}

\hypertarget{incorporating-yearly-variability-in-vital-rates}{%
\subsection{Incorporating yearly variability in vital
rates}\label{incorporating-yearly-variability-in-vital-rates}}

We computed yearly survival probabilities for translocated adults
\(\sigma_{A_T}\) and naturally recruited adults \(\sigma_{A_R}\) from
the posterior distribution of individual state trajectories derived from
mrmr CMR models. Although we observed yearly variability in adult
survival within a population, the magnitude of this variability was
small compared to among-population variability
(Figure 2). Thus, we did not include
yearly within-population variability in adult survival in this analysis.
However, within a population there was substantial yearly variability in
the successful recruitment of adults, greater than what we would expect
from Poisson variability around a mean value. Therefore, we allowed for
yearly variability in the probability of year-1 juvenile survival and
recruitment (additional details provided in \textbf{Estimating model
parameters} below). We also could have included environmental
stochasticity in year-2 juvenile survival and recruitment
\(\sigma_{J_2}\), but our elasticity and sensitivity analysis (see
below, \textbf{Model analysis and simulation}) showed that this
parameter had little effect on host growth rate relative to
\(\sigma_{J_1}\) (Figure~\ref{fig-viability-supp}).

\hypertarget{estimating-model-parameters}{%
\subsection{Estimating model
parameters}\label{estimating-model-parameters}}

The baseline parameter values for the model and how they were estimated
are given in Table~\ref{tbl-param_values}. Parameters \(\sigma_{A_T}\)
and \(\sigma_{A_R}\) were extracted directly from our CMR models (see
\textbf{Materials and Methods - Frog populaton recovery - Estimation of
frog survival and abundance} and \textbf{Supporting Information - Frog
population recovery - CMR model structure} for details). For populations
where we had a sufficient number of PIT-tagged, naturally-recruited
adults, we observed that \(\sigma_{A_T}\) and \(\sigma_{A_R}\) could be
notably different, with \(\sigma_{A_R} > \sigma_{A_T}\)
(Figure~\ref{fig-compare_surv_probs}). For populations lacking
sufficient numbers of naturally-recruited adults, we were unable to
directly estimate \(\sigma_{A_R}\), and instead set
\(\sigma_{A_R} = \sigma_{A_T}\).

To estimate the \(\sigma_{A_R}\), we used the posterior distribution of
predicted true states for naturally-recruited individuals (1=not
recruited, 2=alive, 3=dead, as described in \textbf{Supporting
Information - Frog population recovery - CMR model structure}), then
calculated the posterior probability of individuals surviving between
consecutive primary periods, conditional on being alive in the first
primary period (e.g., given a value of 2 (alive) in the first primary
period, how often was the value still 2 (alive) in the next primary
period compared to 3 (dead)). This yielded posterior distributions for
survival probabilities between primary periods. However, because the
time interval between primary periods differed, the survival
probabilities between different consecutive primary periods were not
directly comparable. To address this, we converted the survival
probabilities between each consecutive primary period to per day death
rates, propagating the uncertainty from the posterior distributions. We
then took a weighted average of these death rates, weighted by the time
interval between primary periods, to get the average per day death rate
over the entire CMR survey. We converted this per day death rate \(d\)
back to a yearly survival probability using
\(\exp(-d \times 365 \text{ days})\). We used the same procedure for
\(\sigma_{A_T}\) such that our estimates of average yearly survival
probability were comparable between \(\sigma_{A_R}\) and
\(\sigma_{A_T}\).

\hypertarget{model-analysis-and-simulation}{%
\subsection{Model analysis and
simulation}\label{model-analysis-and-simulation}}

We performed four analyses on our model. First, we considered a
deterministic version of our model with no yearly heterogeneity in
year-1 juvenile survival and recruitment probability \(\sigma_{J_1}\),
and calculated the predicted long-run growth rate \(\lambda\) of a
population for different values of \(\sigma_{A_R}\) and
\(\sigma_{J_1}\). We then fixed \(\sigma_{J_1} = 0.09\) and calculated
the predicted growth rate of our 12 populations.

Second, we performed a local elasticity and sensitivity analysis on
\(\lambda\) with respect to parameters \(\sigma_{L_1}\),
\(\sigma_{L_2}\), \(\sigma_{L_3}\), \(\sigma_{J_1}\), \(\sigma_{J_2}\),
\(\sigma_{A_R}\), \(F\), \(p_{L_2}\), \(p_{J_1}\) to determine how small
changes in these parameters could influence the long-run deterministic
growth rate of populations (Figure~\ref{fig-viability-supp}). Note
that \(\lambda\) is equally or more sensitive to changes in adult
survival (\(\sigma_{A_R}\)) and survival of year-1 juveniles
(\(\sigma_{J_1}\)) than it is to other model parameters, providing
additional justification for focusing on variation in \(\sigma_{A_R}\)
and \(\sigma_{J_1}\) in our viability analyses.

Third, we defined a version of the model with demographic and
environmental stochasticity, where environmental stochasticity was
represented by among-year variability in \(\sigma_{J_1}\). We used this
model to simulate a one-time introduction of 40 translocated adult
frogs. We ran this simulation 1000 times for each population and
computed the probability of a population becoming extinct after 50 years
given the observed parameter values and environmental stochasticity in
\(\sigma_{J_1}\). We varied the mean recruitment probability
\(\sigma_{J_1}\) from 0 and 0.25 and drew values of \(\sigma_{J_1}\)
each year from a beta distribution with a dispersion parameter of
\(\phi = 2\) (when \(\sigma_{J_1} = 0.5\) and \(\phi = 2\) the beta
distribution is uniform between 0 and 1). Using different values of
\(\phi\) does not qualitatively change the existence of distinct
extinction dynamics between populations with \(\sigma_{A_R} < 0.5\) and
those with \(\sigma_{A_R} > 0.5\). However, increasing yearly
variability in \(\sigma_{J_1}\) increases extinction risk for all
populations. For example, if we set \(\phi = 0.001\), such that in a
given year essentially either all year-1 juveniles survive or all of
them die, populations with \(\sigma_{A_R} > 0.5\) need to have
\(\sigma_{J_1}\) greater than 0.2 to have a 50-year extinction
probability of less 50\%. Because we do not PIT tag juveniles, we do not
have CMR estimates for \(\sigma_{J_1}\) or \(\phi\). However, based on
qualitative and semi-quantitative field observations over 25 years, a
value of \(\sigma_{J_1} = 0.25\) in the presence of Bd is probably a
reasonable estimate for many populations. Thus, we expect our model
predictions to be conservative with regards to population recovery.

Finally, we assessed whether our stochastic model could reproduce
observed trajectories of population recovery. We focused on population
70550 because this was our longest CMR time series for a translocated
population and because this population shows evidence of substantial
post-translocation increases in adult abundance associated with
population establishment and recovery. We simulated our model for 16
years, repeating the simulation 50,000 times. For each run and each
year, we drew \(\sigma_{J_1}\) from a uniform distribution between 0 and
1 (or equivalently a beta distribution with mean 0.5 and \(\phi = 2\)).
Using Approximate Bayesian Computing and rejection sampling
\citep{kosmala2016}, we identified the top 2\% of trajectories (i.e.,
1000 trajectories) that minimized the sum of squared errors between the
observed and predicted data. The yearly \(\sigma_{J_1}\) values
associated with these ``best'' trajectories represented an approximate
posterior distribution \citep{beaumont2010}. Using these best fit
trajectories, we assessed whether our model could qualitatively describe
the patterns of recovery in the observed data for population 70550.

\hypertarget{frog-genome-evolution-2}{%
\section{Frog genome evolution}\label{frog-genome-evolution-2}}

\hypertarget{study-design}{%
\subsection{Study design}\label{study-design}}

To gain insights into the role of evolution in the development of
resistance by MYL frogs, we compared frog exomes sampled in naive versus
recovering populations. Comparing populations with different infection
histories allowed larger sample sizes and replication across the
landscape. The alternative approach of comparing samples from the same
populations before and after Bd exposure isn't feasible in this system
because Bd arrived in most MYL frog populations decades ago and
population persistence/recovery is rare and unpredictable. As a result,
samples from recovering populations collected before and after Bd
exposure are not available and are unlikely to be available in the
future.

Our study design, in which we compared frog genomes in naive and
recovering populations, required sampling populations across the range
of MYL frogs in the southern Sierra Nevada. This is due to the fact that
very few MYL frog populations remain in the naive state, and those that
do are scattered across a wide latitudinal range, from Yosemite National
Park in the north to southern Kings Canyon National Park in south. To
minimize potential confounding effects caused by known variation in frog
genotypes across latitude \citep{byrne2023}, we selected sampling sites
such that both population types were represented across similar
latitudinal ranges (Figure 5).

\hypertarget{go-analysis}{%
\subsection{GO analysis}\label{go-analysis}}

In \textbf{Results-Frog genome evolution}, we describe the stringent set
of outlier variants (identified using a Bonferroni-corrected p-value of
0.01). A liberal set of outlier variants, identified using a
Bonferroni-corrected p-value of 0.05, included 38 outliers (35 SNPs and
3 INDELS) from 30 distinct genes across 16 contigs. We used this liberal
set to determine if any GO biological functions, molecular functions, or
cellular processes were overrepresented. To do this, we retrieved the
BLAST hits and mapped GO terms for each gene in our targeted
transcriptome. We then conducted a statistical overrepresentation test
(Fisher's exact test) using Blast2GO \citep{conesa2005} to compare the
30 unique outlier genes to the complete set of genes in our target
transcriptome. We repeated this process for the set of 35 genes located
in the 9 shared regions of the \emph{F\textsubscript{ST}} and
\(\pi_{diff}\) splined windows.

\hypertarget{genetic-diversity}{%
\subsection{Genetic diversity}\label{genetic-diversity}}

To characterize general patterns of genetic diversity between naive and
recovering populations, we conducted three analyses. First, we
calculated heterozygosity for each sampled frog using VCFtools
\citep{danecek2011}. Second, to characterize genome-wide patterns of
diversity, we used VCFtools to calculate nucleotide diversity (\(\pi\))
in 100kb sliding windows along the genome for each population. Third, we
calculated average \(\pi\) per population within each of the 9 shared
outlier regions identified in the splined window analysis. We compared
average \(\pi\) per population between naive and recovering population
types, and also for each naive and recovering population in Yosemite.
The latter comparison focused on Yosemite because these populations are
the most relevant to the translocated populations described in the
paper.

Average individual-level heterozygosity and genome-wide population-level
\(\pi\) were similar between the naive and recovering groups
(Figure~\ref{fig-violinplot-heterozy};
Figure~\ref{fig-boxplot-genomewide-pi-by-pop}). Within each of the 9
shared outlier regions, \(\pi\) shows considerable variation between
populations and no obvious patterns between the naive and recovering
populations (Figure~\ref{fig-boxplot-pi-by-regionpop}).

\newpage

\newpage

\begin{figure}

{\centering \includegraphics[width=0.60\textwidth]{figures/map_translocation_points.png}

}

\caption{\label{fig-yosemap}Map showing the locations of translocated
and donor MYL frog populations in Yosemite National Park (park boundary
indicated by gray polygon). Symbol shapes indicate the donor population
used for each translocation site, and 5-digit numbers identify each
donor and translocation site. To obscure the exact locations of
populations, random noise was added to all point coordinates. Inset map
shows the location of Yosemite within California. In both maps,
elevation is indicated by the colored hillshade layer (dark green =
lowest elevation, white = highest elevation).}

\end{figure}\clearpage

\newpage

\begin{figure}

{\centering \includegraphics[width=0.80\textwidth]{figures/bdload_beforeafter.png}

}

\caption{\label{fig-bdload-beforeafter}For frogs translocated to each of
the 12 recipient sites, Bd loads for the period immediately prior to
translocation versus during the 1-year period after translocation. Bd
loads are expressed as the number of ITS1 copies per skin swab, as
estimated by qPCR of the Bd ITS1 region. Box plots show medians, first
and third quartiles, largest and smallest values within 1.5x
interquartile range, and values outside the 1.5x interquartile range.
Samples sizes are provided immediately above the x-axis. The red dashed
horizontal line indicates the approximate load above which severe
disease is typically observed \citep{vredenburg2010}. This threshold
(5.8 ITS1 copies - on a log\textsubscript{10} scale) is commonly
exceeded during Bd epizootics, but is exceeded only rarely in recovering
populations \citep{briggs2010}.}

\end{figure}\clearpage

\newpage

\begin{figure}

{\centering \includegraphics[width=0.60\textwidth]{figures/mcmc_areas_m1d.png}

}

\caption{\label{fig-survival-postdens}Results from the among-site
meta-analysis, showing that Bd load is not an important predictor of
post-translocation frog survival. Depicted distributions are the
estimated posterior density curves and shaded 95\% uncertainty intervals
for the intercept and all predictor variables from the best model. In
the Bayesian framework in which the model was developed, variables are
considered important predictors if the associated uncertainty interval
does not overlap zero (indicated by the dashed red line). Predictor
variables shown on the y-axis are defined as follows: length = frog
size, snow\_t = winter severity in the year of translocation (measured
on April 1), snow\_t1 = winter severity in the year following
translocation (measured on April 1), day = day of year on which a
translocation was conducted, bd\_load = Bd load, elevation = site
elevation, order\_first = within-site translocation order, sex\_male =
frog sex, donor\_70567 and donor\_72996 = donor population.}

\end{figure}\clearpage

\newpage

\begin{figure}

{\centering \includegraphics[width=0.60\textwidth]{figures/compare_surv_probs.jpg}

}

\caption{\label{fig-compare_surv_probs}Comparison of average yearly
adult survival probabilities for adults translocated to each of 3 sites
versus adults that were naturally recruited at each site (as a result of
reproduction by translocated frogs). In contrast to
Figure 2, these are not survival
probabilities from the first year following translocation, but instead
represent averaged survival probabilities across multiple years and
cohorts. Points are median estimates and error bars give the 95\%
uncertainty intervals around the estimates, accounting for yearly
variation in survival probabilities. All estimates were derived using
the mrmr package, and the methods for calculation are described in
\textbf{Supporting Information - Population viability modeling -
Estimating model parameters}.}

\end{figure}\clearpage

\newpage

\begin{figure}

{\centering \includegraphics[width=0.95\textwidth]{figures/pop_viability_figures_for_supp.jpg}

}

\caption{\label{fig-viability-supp}Sensitivity analysis of the
stage-structured MYL frog model. A. Elasticity of \(\lambda\) with
changes in nine parameters: \(\sigma_{A_R}\) (yearly survival
probability of naturally recruited adults), \(F\) (number of eggs
produced by a female frog in a year that successfully hatch),
\(\sigma_{J_1}\) (yearly probability of survival of year-1 juveniles
that also affects recruitment), \(\sigma_{J_2}\) (yearly probability of
survival and recruitment of year-2 juveniles), \(p_{L_2}\) (probability
that year-2 tadpoles metamorph into juveniles), \(p_{J_1}\) (probability
that year-1 juveniles mature to adults), \(\sigma_{L_1}\) (survival
probability of year-1 tadpoles), \(\sigma_{L_2}\) (survival probability
of year-2 tadpoles), and \(\sigma_{L_3}\) (survival probability of
year-3 tadpoles). B. Sensitivity of \(\lambda\) with changes in the same
nine parameters. Note that \(F\) has different units than the other
parameters, which is why its sensitivity is low. Sensitivities of all
other parameters are comparable. We did not include \(p_{L_1}\) as this
parameter is 1 across the populations in this study (Table S1) --
tadpoles spend at least a year as tadpoles before metamorphosing.
Elasticity and sensitivity is calculated at the default parameter values
for each population and \(\sigma_{J_1}\) = 0.09.}

\end{figure}\clearpage

\newpage

\begin{figure}

{\centering \includegraphics[width=0.85\textwidth]{figures/pca_qq_manhattan.png}

}

\caption{\label{fig-selectionresults}Results from the analysis of
individual variants, showing possible signatures of selection in
recovering MYL frog populations. (A) PCA calculated from binary SNPs
showing the genomic relationship of samples. Numeric labels and colors
match those in Figure 5. Populations 1-7 are
\emph{R. sierrae} and populations 8 and 9 are \emph{R. muscosa}. (B)
qqplot showing observed and expected p-values for 148,307 SNPs and
INDELS. Dashed line shows p-value that identifies outliers. (C)
Manhattan plot showing p-value for each SNP. SNPs are sorted by genomic
position and contigs are sorted by size. Red line shows p-value that
identifies outliers. Outlier SNPs above this threshold are highlighted
and labeled. Bold labels indicate the presence of at least one
non-synonymous SNP in that gene.}

\end{figure}\clearpage

\newpage

\begin{figure}

{\centering \includegraphics[width=0.95\textwidth]{figures/synteny_figure.png}

}

\caption{\label{fig-synteny-plot}Synteny plot showing conserved gene
order in \emph{Xenopus tropicalis} and \emph{Rana muscosa} for the
outlier region containing MHC Class I Classical and MHC Class III gene
regions. The plot was created with SimpleSynteny \citep{veltri2016}
using \emph{Xenopus tropicalis} Chromosome 8 (NC\_030684.2, genbank
accession GCA\_000004195.4) and \emph{Rana muscosa} Contig19. Asterisks
indicate the location of SNP outliers in the TCF19 and VARS1 genes. Gap
sizes for each contig representation are labeled.}

\end{figure}\clearpage

\newpage

\begin{figure}

{\centering \includegraphics[width=0.85\textwidth]{figures/ave_pi_per_popn_yose.png}

}

\caption{\label{fig-ave-pi-per-popn-yose}For each study population
located in Yosemite National Park (populations 1-4 shown in the map),
\(\pi\) for the 9 shared outlier regions identified in the splined
window analysis. Populations 1 and 4 were the source of frogs for 20 of
the 24 frog translocations, and the donor population for the remaining
four translocations is located adjacent to populations 3 and 4 (see also
Figure~\ref{fig-yosemap}). Bars are color-coded by population type
(naive, recovering). The corresponding gene annotations for each region
are as follows (gene regions described in the \textbf{Results} and
\textbf{Discussion} are indicated with an asterisk): (1) C6*, C7*, (2)
DDX10, ZBTB24, (3) SULTR-like, TRAF3IP2, VGLL3, EXOC1, (4) FLOT1, TUBB,
MDC1, CCHCR1, TCF19*, HSP70, LSM2, VARS1*, (5) GCC2, CFAP251, PEG10, (6)
ERO1A, GVINP1*, (7) PPP1R12A, TSPAN4, PAWR, MFRP, MAX, PPP6R3, (8)
C6H5ORF22, PKS6, BSPRY, MPV17, and (9) CAD, ATRAID, GPN1. Region 4 is
the extended MHC region that includes two SNPs identified by the GEMMA
analysis in genes TCF19 and VARS1. In the map, the four populations
included in panels 1-9 are color-coded by population type, and dashed
lines delineate the genetic clusters as defined in
Figure~\ref{fig-selectionresults}A.}

\end{figure}\clearpage

\newpage

\begin{figure}

{\centering \includegraphics[width=0.60\textwidth]{figures/violin_plot_heterozy_by_group.png}

}

\caption{\label{fig-violinplot-heterozy}Violin plots showing individual
heterozygosity for the Bd-naive and recovering populations included in
the frog evolution study. Individual data points are represented by
their corresponding site number (from
Figure 5).}

\end{figure}\clearpage

\newpage

\begin{figure}

{\centering \includegraphics[width=0.70\textwidth]{figures/boxplot_genomewide_pi_by_pop.png}

}

\caption{\label{fig-boxplot-genomewide-pi-by-pop}For each population in
the frog evolution study (Figure 5), boxplots
showing nucleotide diversity (\(\pi\)) within 100kb sliding windows
along the genome. The y-axis has been log\textsubscript{10}-transformed
for display purposes. Sites are color-coded as ``Bd-naive'' or
``recovering'', and vertical dashed lines show the genetic breaks
between frog populations as described in
Figure~\ref{fig-selectionresults}A).}

\end{figure}\clearpage

\newpage

\begin{figure}

{\centering \includegraphics[width=0.70\textwidth]{figures/boxplot_pi_by_regionpop.png}

}

\caption{\label{fig-boxplot-pi-by-regionpop}For each of the 9 shared
outlier regions identified by the splined window analysis, boxplots
showing \(\pi\) values for naive and recovering populations. The
corresponding gene annotations for each region are as follows (gene
regions described in the \textbf{Results} and \textbf{Discussion} are
indicated with an asterisk): (1) C6*, C7*, (2) DDX10, ZBTB24, (3)
SULTR-like, TRAF3IP2, VGLL3, EXOC1, (4) FLOT1, TUBB, MDC1, CCHCR1,
TCF19*, HSP70, LSM2, VARS1*, (5) GCC2, CFAP251, PEG10, (6) ERO1A,
GVINP1*, (7) PPP1R12A, TSPAN4, PAWR, MFRP, MAX, PPP6R3, (8) C6H5ORF22,
PKS6, BSPRY, MPV17, and (9) CAD, ATRAID, GPN1.}

\end{figure}\clearpage

\newpage

\hypertarget{tbl-param_values}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3611}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2639}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3750}}@{}}
\caption{\label{tbl-param_values}Description and values of parameters
used in the model. All survival probabilities are in the presence of the
fungal pathogen Bd.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Parameter}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Value}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Source}
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Parameter}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Value}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Source}
\end{minipage} \\
\midrule()
\endhead
\(\sigma_{L_1}\), Yearly survival probability of year-1 tadpoles & 0.7 &
Estimated from field data, observations, natural history knowledge \\
\(\sigma_{L_2}\), Yearly survival probability of year-2 tadpoles & 0.7 &
Estimated from field data, observations, natural history knowledge \\
\(\sigma_{L_3}\), Yearly survival probability of year-3 tadpoles & 0.7 &
Estimated from field data, observations, natural history knowledge \\
\(\sigma_{J_1}\), Yearly survival probability of year-1 juveniles &
Varies yearly & Varies. Bounds estimated from field data, observations,
natural history knowledge \\
\(\sigma_{J_2}\), Yearly survival probability of year-2 juveniles & 0.5
& Estimated from field data, observations, natural history knowledge \\
\(\sigma_{A_R}\), Yearly survival probability of naturally recruited
adults & Varies by population & Estimated from CMR studies \\
\(\sigma_{A_T}\), Yearly survival probability of translocated adults &
Varies by population & Estimated from CMR studies \\
\(p_{L_1}\), Probability of a year-1 tadpoles remaining as a tadpoles &
1 & Estimated from field data, observations, natural history
knowledge \\
\(p_{L_2}\), Probability of a year-2 tadpoles remaining as a tadpoles &
0.25 & Estimated from field data, observations, natural history
knowledge \\
\(p_{J_1}\), Probability of a year-1 juvenile remaining as a juvenile &
0.25 & Estimated from field data, observations, natural history
knowledge \\
\(p_F\), Probability of a adult female reproducing in a year & 0.5 &
Could be as high at 1, based on field observations \\
\(F\), Number of surviving eggs produced by an adult female & 100 & From
observations of captive frogs \\
\bottomrule()
\end{longtable}

\newpage

%%% Add this line AFTER all your figures and tables
\FloatBarrier

\dataset{gemma\_outliers\_all.csv}{Set of liberal SNP and INDEL outlier variants (Bonferroni corrected p-value $<$ 0.05), identified via GEMMA.}

\dataset{gemma\_outliers\_strict\_freq.csv}{Set of strict SNP and INDEL outlier variants (Bonferroni corrected p-value $<$ 0.01), identified via GEMMA.
Additional information includes variant location within the gene (predicted\_gene\_loc) and whether the variant is synonymous or nonsynonymous (predicted\_effect\_AA).}

\dataset{spline\_window\_shared\_outliers.csv}{Description of overlapping \emph{F\textsubscript{ST}} and
\(\pi_{diff}\) outlier windows, as identified in the splined window
analysis.}

\dataset{spline\_window\_gene\_details.csv}{Annotation information for all genes within each of the
overlapping outlier windows in dataset S3}

\bibliography{translocation}

\end{document}